---
title: "figures"
author: "Maxwell C. Cook"
output: html_document
---

```{r setup, include=FALSE, echo=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("setup.R")
```

```{r}
# Grab a glimpse of the data
glimpse(ics.conus)
```

Filter out incidents with 0 acres or with missing temporal information.

```{r}
# Remove records reporting 0 acres (likely erroneous)
ics.conus <- ics.conus %>% 
  filter(FINAL_ACRES != 0, FINAL_KM2 != 0,
         !is.na(DISCOVERY_DATE))
```

Load the joined database.

```{r message=F, warning=F}
ics.fired <- st_read("../../data/spatial/mod/ics-fired/ics-fired_west_plus_2001to2020.gpkg")
```

Print some summary statistics for CONUS and West+.

```{r}
# Burned area
print(paste("Burned Area (acres): ", format(round(as.numeric(sum(ics.conus$FINAL_ACRES)), 1), 
                                            nsmall=1, big.mark=","), sep=""))
# Homes
print(paste("Homes Destroyed: ", round(sum(ics.conus$STR_DESTROYED_RES_TOTAL), 0), sep=""))
# Cost
tmp <- ics.conus %>% filter(!is.na(PROJECTED_FINAL_IM_COST))
print(paste("Projected Suppression Costs: ", format(round(as.numeric(sum(tmp$PROJECTED_FINAL_IM_COST)), 1), 
                                            nsmall=1, big.mark=","), sep=""))
## For west+
print("~~~~~~~~West-Wide~~~~~~~~")
west.plus <- c("AZ", "CO", "NV", "WY", "CA", "ID", "WA", "OR", "NM", "MT", "UT", "TX", "OK")
ics.west.plus <- ics.conus %>% filter(STUSPS %in% west.plus)
# Burned area
print(paste("Burned Area (acres): ", format(round(as.numeric(sum(ics.west.plus$FINAL_ACRES)), 1), 
                                            nsmall=1, big.mark=","), sep=""))
# Homes
print(paste("Homes Destroyed: ", round(sum(ics.west.plus$STR_DESTROYED_RES_TOTAL), 0), sep=""))
# Cost
tmp2 <- ics.west.plus %>% filter(!is.na(PROJECTED_FINAL_IM_COST))
print(paste("Projected Suppression Costs: ", format(round(as.numeric(sum(tmp2$PROJECTED_FINAL_IM_COST)), 1), 
                                            nsmall=1, big.mark=","), sep=""))
## West-wide percentages
sum(tmp2$PROJECTED_FINAL_IM_COST) / sum(tmp$PROJECTED_FINAL_IM_COST) * 100
rm(tmp)
rm(tmp2)
```

Print some summaries for FIRED.

```{r}
sum(fired.conus$tot_ar_km2)
```

Plot burned area reported by the ICS-209-PLUS West-wide and CONUS.

```{r fig.width=5, fig.height=3}
# Group the western U.S. incidents
ics.west.grp <- ics.west %>% 
  group_by(START_YEAR) %>%
  summarize(burned_area = sum(FINAL_ACRES)) %>%
  ungroup()
# Grab the top 5
top5 <- ics.west.grp %>% slice_max(burned_area, n=5)
# Plot
f1 <- ics.conus %>% group_by(START_YEAR) %>%
  summarize(burned_area = sum(FINAL_ACRES)) %>%
  ungroup() %>%
  ggplot(aes(x=START_YEAR, y=burned_area)) +
  geom_line(data=ics.west.grp, aes(x=START_YEAR, y=burned_area, color="Western U.S."), position="stack",
            linetype = 6, size = 0.8) +
  geom_point(data=ics.west.grp, aes(x=START_YEAR, y=burned_area), 
             shape=21, size = 1.75, color = "gray20", fill = "gray50") +
  geom_text(data=. %>% filter(START_YEAR %in% top5$START_YEAR), 
            aes(label=paste(round(burned_area / 1e6, 1), "M")),
            position=position_nudge(x=-1.2,y=-10), size=3) +
  geom_line() +
  geom_point(shape=21, size = 2.5, color = "gray20", fill = "red") +
  scale_y_continuous(labels = scales::label_number(suffix = " M", scale = 1e-6)) +
  scale_x_continuous(limits=c(1999, 2020)) +
  scale_color_manual(name = "", values=c("Western U.S."="gray60")) +
  labs(x="\nIgnition Year", y="Burned Area (acres)\n", title="Conterminous U.S. burned area (1999-2020)\n") +
  theme_minimal() +
  theme(plot.title = element_text(size = 12, hjust=-0.4),
        legend.position=c(0.2, 1.1))
f1
ggsave(f1, file = "../../figs/BurnedArea_1999to2020.png",
       width=5, height=3, dpi = 300) # adjust dpi accordingly
```

Plot structures destroyed by incident.

```{r fig.height=3.5, fig.width=6}
# Grab the top 15
top15 <- ics.west.plus %>% slice_max(STR_DESTROYED_RES_TOTAL, n=15)
# Plot and label
f2 <- ggplot(data=ics.west.plus, aes(x=START_YEAR, y=STR_DESTROYED_RES_TOTAL, group=factor(INCIDENT_NAME), 
                          fill=STR_DESTROYED_RES_TOTAL), color="grey10") +
  geom_bar(stat="identity", position='stack') +
  scale_fill_viridis_c(option="plasma", trans="sqrt") +
  labs(x="Ignition Year\n", y="Homes Destroyed\n", fill="",
       title="Homes destroyed by Western wildfires (1999-2020)\n",
       caption="Maxwell C. Cook, PhD Student, Department of Geography\nData source: St. Denis et al., 2020 (updated)") +
  ggrepel::geom_text_repel(data=. %>% filter(INCIDENT_ID %in% top15$INCIDENT_ID), 
                           aes(label=INCIDENT_NAME), position=position_jitterdodge(0.5), size = 2) +
  theme_minimal() +
    # coord_cartesian(ylim=c(0, 1150)) +
  theme(plot.title = element_text(size = 12),
        plot.subtitle = element_text(size=8),
        plot.caption = element_text(size=7),
        axis.title.x = element_text(size=9),
        axis.title.y = element_text(size=9),
        axis.text.y = element_text(size=8),
        axis.text.x = element_text(size=8),
        legend.position="none")
f2
rm(top15)
ggsave(f2, file = "../../figs/Westwide_HomeLoss_1999to2020.png",
       width=6, height=3.5, dpi = 300) # adjust dpi accordingly
```

Plot cumulative residential structures destroyed from ICS-209-PLUS for CONUS and West.

```{r warning=F, message=F, fig.width=5, fig.height=3}
# Get some text
sum(ics.west$STR_DESTROYED_RES_TOTAL)
sum(ics.conus$STR_DESTROYED_RES_TOTAL)
## Group sum ics.west
ics.west.grp <- ics.west %>% 
  group_by(START_YEAR) %>%
  summarize(destr_res = sum(STR_DESTROYED_RES_TOTAL)) %>%
  mutate(destr_res_c = cumsum(destr_res))
# plot
f3 <- ics.conus %>% as_tibble() %>%
  group_by(START_YEAR) %>%
  summarize(destr_res = sum(STR_DESTROYED_RES_TOTAL)) %>%
  mutate(destr_res_c = cumsum(destr_res)) %>%
  ungroup() %>%
  ggplot() +
  geom_area(aes(x=START_YEAR, y=destr_res_c), position='stack', fill="grey60") +
  geom_line(data=ics.west.grp, aes(x=START_YEAR, y=destr_res_c, color="Western U.S."), position="stack",
            linetype = 4, size = 1.05) +
  scale_color_manual(name = "", values=c("Western U.S."="black")) +
  scale_y_continuous(labels = scales::label_number(suffix = " K", scale = 1e-3)) + 
  labs(x="Ignition Year", y="Structures Destroyed",
       title="Homes destroyed by wildfire (1999-2020)\n")+
  # annotation_custom(grob) +
  theme_minimal() +
  theme(legend.position=c(0.2, 1.1))
f3
ggsave(f3, file = "../../figs/Cumulative_HomeLoss_1999to2020.png",
       width=5, height=3, dpi = 300) # adjust dpi accordingly
```

Map structures destroyed and burned area.

```{r fig.height=5, fig.width=5}
map1 <- ggplot() +
  geom_sf(data=ics.conus%>%filter(STR_DESTROYED_RES_TOTAL==0), color="gray60", size=0.3, alpha=0.9) +
  geom_sf(data=states%>%filter(STUSPS %in% states.sub), fill=NA, color="gray20") +
  geom_sf(data=ics.conus%>%filter(STR_DESTROYED_RES_TOTAL>0)%>%arrange(STR_DESTROYED_RES_TOTAL),
          aes(color=STR_DESTROYED_RES_TOTAL, size=FINAL_ACRES), alpha=0.8) +
  scale_size(range = c(1,6)) +
  scale_color_viridis_c(option="inferno", trans="log10") +
  labs(color = "Homes Destroyed") +
  coord_sf(crs = st_crs(ics.conus)) +
  theme_void() +
  guides(color = guide_colourbar(position="top", barwidth = 0.4, barheight = 6, ticks=F,
                                 label.position = "right", title.position = "left",
                                 label.theme = element_text(angle = 0, size = 8)),
         size="none") +
  theme(legend.title = element_text(angle = 90, size=10),
        legend.position=c(0.85, 0.7))
map1
ggsave(map1, file = paste("../../figs/Map1-ICS209_HomesDestroyed_Size_CONUS.png", sep=""), 
       width=5, height=5, dpi = 350) # adjust dpi accordingly
```
Plot cumulative structures destroyed from ICS-209-PLUS.

```{r warning=F, message=F, fig.width=5, fig.height=3}
# Create some text grob
# Text
grob <- grobTree(
   textGrob(paste("Total Destroyed Structures: ", as.integer(sum(ics$STR_DESTROYED_TOTAL)), sep=""), 
            x=0.05,  y=0.90, hjust=0, gp=gpar(col="grey20", fontsize=8, fontface="italic")),
   textGrob(paste("Destroyed Residential Structures: ", as.integer(sum(ics$STR_DESTROYED_RES_TOTAL)), sep=""), 
            x=0.05,  y=0.80, hjust=0, gp=gpar(col="grey20", fontsize=8, fontface="italic")),
   textGrob(paste("Number of 'Destructive' Incidents: ", dim(ics %>% filter(STR_DESTROYED_TOTAL>=1))[1], sep=""), 
            x=0.05,  y=0.70, hjust=0, gp=gpar(col="grey20", fontsize=8, fontface="italic")),
   textGrob(paste("Number of 'Destructive' Incidents (Res.): ", dim(ics %>% filter(STR_DESTROYED_RES_TOTAL>=1))[1], sep=""), 
            x=0.05,  y=0.60, hjust=0, gp=gpar(col="grey20", fontsize=8, fontface="italic")))
# plot
f1 <- ics %>% as_tibble() %>%
  group_by(START_YEAR) %>%
  summarize(destr_tot = sum(STR_DESTROYED_TOTAL),
            destr_res = sum(STR_DESTROYED_RES_TOTAL)) %>%
  mutate(destr_tot_c = cumsum(destr_tot),
         destr_res_c = cumsum(destr_res)) %>%
  ungroup() %>%
  ggplot() +
    geom_area(aes(x=START_YEAR, y=destr_tot_c), fill="#8c2d04") +
    geom_area(aes(x=START_YEAR, y=destr_res_c), position='identity', fill="#fee391") +
  scale_y_continuous(labels = scales::label_number(suffix = " K", scale = 1e-3)) + 
  geom_bar(aes(x=START_YEAR, y=destr_tot), position='identity', stat='identity') +
  labs(x="Ignition Year", y="Structures Destroyed", fill="Legend")+
  annotation_custom(grob) +
  theme_minimal()
f1
ggsave(f1, file = paste("C:/Users/mccoo/OneDrive/mcook/ics209/", "SuppFigure_StructureLoss.png", sep=""), 
       width=5, height=3, dpi = 700) # adjust dpi accordingly
```

Isolate CO fires.

```{r}
ics.co <- ics.west %>% filter(STUSPS == 'CO')
paste("Number of wildfire incidents in 209s: ", dim(ics.co)[1])
paste("Number of destructive (n>1) CO incidents: ", dim(ics.co%>%filter(STR_DESTROYED_RES_TOTAL>0))[1])
paste("Total number of structures lost: ", sum(ics.co$STR_DESTROYED_TOTAL)+991)
paste("Total number of res. structures lost: ", sum(ics.co$STR_DESTROYED_RES_TOTAL)+1084)
tmp <- ics.co %>% filter(START_YEAR <= 2010)
paste("Res. structures lost 1999-2010: ", sum(tmp$STR_DESTROYED_RES_TOTAL))
tmp <- ics.co %>% filter(START_YEAR > 2010)
paste("Res. structures lost 2011-2021: ", sum(tmp$STR_DESTROYED_RES_TOTAL)+1084)
# Marshall Fire
paste("Marshall Fire % Destruction: ", 1091/sum(ics.co$STR_DESTROYED_TOTAL)*100)
paste("Marshall Fire % Destruction (residential): ", 1084/sum(ics.co$STR_DESTROYED_RES_TOTAL)*100)
```

Plot CO destruction.

```{r warning=F, message=F, fig.width=5.25, fig.height=3}
# plot
f2 <- ics.co %>% as_tibble() %>%
  group_by(START_YEAR) %>%
  summarize(destr_tot = sum(STR_DESTROYED_TOTAL),
            destr_res = sum(STR_DESTROYED_RES_TOTAL)) %>%
  mutate(destr_tot_c = cumsum(destr_tot),
         destr_res_c = cumsum(destr_res)) %>%
  ungroup() %>%
  add_row(START_YEAR = 2021, destr_tot = 1091, destr_res = 1084, destr_tot_c = 5357, destr_res_c = 3734) %>%
  ggplot() +
    geom_area(aes(x=START_YEAR, y=destr_tot_c, fill="Total Structures (Cumulative)")) +
    geom_area(aes(x=START_YEAR, y=destr_res_c, fill="Residential Structures (Cumulative)"), position='identity') +
    # scale_y_continuous(labels = scales::label_number(suffix = " K", scale = 1e-3)) +
    geom_bar(aes(x=START_YEAR, y=destr_tot, fill="Annual Loss (Total)"), stat='identity') +
    scale_fill_manual(values = c("grey15", "#8c2d04", "#fee391"), guide = guide_legend(reverse = TRUE)) +
    # geom_bar(aes(x=START_YEAR, y=destr_res, fill="Res. Structures"), position='identity', stat='identity') +
    labs(x="\nIgnition Year", y="Structures Destroyed\n", fill="",
         title="Colorado Wildfire-Related Structure Loss (1999-2021)\n") +
    theme_minimal() +
    theme(plot.title = element_text(size = 12), legend.position=c(0.35, 0.8))
f2
ggsave(f2, file = paste("C:/Users/mccoo/OneDrive/mcook/ics209/figs/", "CO_StructureLoss.png", sep=""), 
       width=5, height=3, dpi = 700) # adjust dpi accordingly
```

Plot CO destruction. Not cumulative.

```{r warning=F, message=F, fig.width=5.25, fig.height=3}
# plot
f3 <- ics.co %>% as_tibble() %>%
  group_by(START_YEAR) %>%
  summarize(destr_tot = sum(STR_DESTROYED_TOTAL),
            destr_res = sum(STR_DESTROYED_RES_TOTAL)) %>%
  ungroup() %>%
  add_row(START_YEAR = 2021, destr_tot = 1091, destr_res = 1084) %>%
  ggplot() +
    geom_bar(aes(x=START_YEAR, y=destr_res), fill="#a63603", stat="identity", position='identity') +
    labs(x="\nIgnition Year", y="Homes Destroyed\n", fill="",
         title="Colorado Wildfire-Related Home Loss (1999-2021)\n") +
    theme_minimal() +
    theme(plot.title = element_text(size = 12), legend.position="top")
f3
ggsave(f3, file = paste("C:/Users/mccoo/OneDrive/mcook/ics209/figs/", "CO_StructureLoss_NoCSum_RES.png", sep=""), 
       width=5, height=3, dpi = 700) # adjust dpi accordingly
```

Plot CO destruction. Not cumulative. By incident.

```{r warning=F, message=F, fig.width=5.5, fig.height=3.5}
plasma_pal <- viridis::plasma(6)
# plot
f4 <- ics.co %>% as_tibble() %>%
  add_row(INCIDENT_NAME = "MARSHALL FIRE", START_YEAR = 2021, STR_DESTROYED_TOTAL = 1091, STR_DESTROYED_RES_TOTAL = 1084) %>%
  ggplot(aes(x=START_YEAR, y=STR_DESTROYED_RES_TOTAL, group=factor(INCIDENT_NAME), fill=STR_DESTROYED_RES_TOTAL), color="grey10") +
    geom_bar(stat="identity", position='stack') +
    scale_fill_gradientn(colors = viridis_pal(begin=0, end=0.9, option="plasma")(3)) +
    labs(x="Ignition Year\n", y="Homes Destroyed\n", fill="",
         title="Homes destroyed by Colorado wildfires (1999-2021)\n",
         caption="Maxwell C. Cook, PhD Student, Department of Geography,\nDr. Jennifer K. Balch, Director, Earth Lab\nData source: St. Denis et al., 2020 (updated)") +
    theme_minimal() +
    coord_cartesian(ylim=c(0, 1150)) +
    theme(plot.title = element_text(size = 11.5, hjust = -0.55),
          plot.subtitle = element_text(size=8),
          plot.caption = element_text(size=7),
          axis.title.x = element_text(size=9),
          axis.title.y = element_text(size=9),
          axis.text.y = element_text(size=8),
          axis.text.x = element_text(size=8),
          legend.position="none")
f4
ggsave(f4, file = paste("C:/Users/mccoo/OneDrive/mcook/ics209/figs/", "CO_StructureLoss_NoCSum_RES_byIncident.png", sep=""), 
       width=5.5, height=3.5, dpi = 700) # adjust dpi accordingly
```

Show top 15 most destructive fires in western U.S.

```{r}
ics.top <- top_n(ics, 15, STR_DESTROYED_RES_TOTAL)
```

```{r warning=F, message=F, fig.width=5.25, fig.height=3}
# plot
f4 <- ics.co %>% as_tibble() %>%
  group_by(START_YEAR) %>%
  summarize(destr_tot = sum(STR_DESTROYED_TOTAL),
            destr_res = sum(STR_DESTROYED_RES_TOTAL)) %>%
  mutate(destr_tot_c = cumsum(destr_tot),
         destr_res_c = cumsum(destr_res)) %>%
  ungroup() %>%
  add_row(START_YEAR = 2021, destr_tot = 991, destr_res = 969, destr_tot_c = 5357, destr_res_c = 3734) %>%
  ggplot() +
    geom_area(aes(x=START_YEAR, y=destr_res_c, fill="Residential Structures (Cumulative)"), position='identity') +
    # scale_y_continuous(labels = scales::label_number(suffix = " K", scale = 1e-3)) +
    geom_bar(aes(x=START_YEAR, y=destr_res, fill="Annual Loss (Total)"), stat='identity') +
    scale_fill_manual(values = c("grey15", "#a63603"), guide = guide_legend(reverse = TRUE)) +
    # geom_bar(aes(x=START_YEAR, y=destr_res, fill="Res. Structures"), position='identity', stat='identity') +
    labs(x="\nIgnition Year", y="Homes Destroyed\n", fill="",
         title="Colorado Wildfire-Related Home Loss (1999-2021)\n") +
    theme_minimal() +
    theme(plot.title = element_text(size = 12), legend.position=c(0.35, 0.8))
f4
ggsave(f4, file = paste("C:/Users/mccoo/OneDrive/mcook/ics209/figs/", "CO_StructureLoss_RES.png", sep=""), 
       width=5, height=3, dpi = 700) # adjust dpi accordingly
```

Timing of destruction based on ignition date.

```{r}
ics.co %>%
  filter(STR_DESTROYED_RES_TOTAL>0) %>%
  group_by(START_MONTH) %>%
  summarize(n = n()) %>%
  glimpse()
print("____________________________")
ics.co %>%
  group_by(START_MONTH) %>%
  summarize(n = n()) %>%
  glimpse()
```

```{r}
tmp <- ics.co %>% group_by(START_YEAR) %>%
  mutate(FINAL_HA = FINAL_ACRES*0.404686) %>%
  summarize(burned_area = sum(FINAL_ACRES),
            str_des = sum(STR_DESTROYED_RES_TOTAL)) %>%
  ungroup() %>%
  add_row(START_YEAR = 2021, burned_area = 132.97, str_des = 991) %>%
  mutate(rate = str_des / burned_area)
glimpse(tmp)
```

Homes per acres burned area by year. 2021 = ~32,860 acres (132.97)

```{r warning=F, message=F, fig.width=5.25, fig.height=3}
f3 <- ics.co %>% group_by(START_YEAR) %>%
  mutate(FINAL_HA = FINAL_ACRES*0.404686) %>%
  summarize(burned_area = sum(FINAL_KM2),
            str_des = sum(STR_DESTROYED_RES_TOTAL)) %>%
  ungroup() %>%
  add_row(START_YEAR = 2021, burned_area = 132.97, str_des = 991) %>%
  mutate(rate = str_des / burned_area) %>%
  ggplot() +
  geom_bar(aes(x=START_YEAR, y=rate), stat="identity") +
  geom_line(aes(x=START_YEAR, y=log(burned_area), color="log(Burned Area) (KM2)")) +
  labs(x="\nIgnition Year", y="Structures per Km2 Burned\n", color="",
         title="Colorado Wildfire-Related Structure Loss Rate (1999-2021)\n") +
  theme_minimal() +
  theme(plot.title = element_text(size = 11), legend.position=c(0.2, 1.1))
f3
ggsave(f3, file = paste("C:/Users/mccoo/OneDrive/mcook/ics209/figs/", "CO_StructureLoss_Rate.png", sep=""), 
       width=5, height=3, dpi = 700) # adjust dpi accordingly
```

MTBS record burned area (and burned area December).

```{r}
co <- st_read("C:/Users/mccoo/OneDrive/mcook/data/boundaries/political/colorado.gpkg", quiet=T) %>%
  st_transform(st_crs(mtbs))
glimpse(mtbs)
mtbs.co <- mtbs %>% st_intersection(., co) %>% mutate(MTBS_YEAR = as.integer(MTBS_YEAR))
```

```{r}
ics.co.20 <- ics.co %>% filter(START_YEAR == 2020)
sum(ics.co.20$FINAL_ACRES)
```

```{r warning=F, message=F, fig.width=5.25, fig.height=3}
f6 <- mtbs.co %>% group_by(MTBS_YEAR) %>%
  summarize(burned_area = sum(MTBS_ACRES)) %>%
  ungroup() %>%
  add_row(MTBS_YEAR=2020, burned_area=678721.3) %>%
  add_row(MTBS_YEAR=2021, burned_area=32860) %>%
  ggplot(aes(x=MTBS_YEAR, y=burned_area)) +
  geom_line() +
  geom_point(shape=21, size = 3, color = "gray30", fill = "red") +
  scale_y_continuous(labels = scales::label_number(suffix = " K", scale = 1e-3)) +
  scale_x_continuous(limits=c(1984, 2021)) +
  labs(x="\nIgnition Year", y="Burned Area (acres)\n", title="Colorado Burned Area (1984-2021)\n") +
  theme_minimal() +
  theme(plot.title = element_text(size = 14))
f6
ggsave(f6, file = paste("C:/Users/mccoo/OneDrive/mcook/ics209/figs/", "CO_BurnedArea_1984to2021.png", sep=""), 
       width=5, height=3, dpi = 700) # adjust dpi accordingly
```

```{r warning=F, message=F, fig.width=5.25, fig.height=3}
dec <- ics.co %>% filter(START_MONTH=="Dec")
f9 <- mtbs.co %>% group_by(MTBS_YEAR) %>%
  summarize(burned_area = sum(MTBS_ACRES)) %>%
  ungroup() %>%
  add_row(MTBS_YEAR=2020, burned_area=678721.3) %>%
  add_row(MTBS_YEAR=2021, burned_area=32860) %>%
  ggplot(aes(x=MTBS_YEAR, y=log(burned_area))) +
  geom_line() +
  geom_point(shape=21, size = 3, color = "gray30", fill = "red") +
  scale_y_continuous(labels = scales::label_number(suffix = " K", scale = 1e-3)) +
  scale_x_continuous(limits=c(1984, 2021)) +
  labs(x="\nIgnition Year", y="Burned Area (acres)\n", title="Colorado Burned Area (1984-2021)\n") +
  theme_minimal() +
  theme(plot.title = element_text(size = 14))
f9
ggsave(f6, file = paste("C:/Users/mccoo/OneDrive/mcook/ics209/figs/", "CO_BurnedArea_1984to2021.png", sep=""), 
       width=5, height=3, dpi = 700) # adjust dpi accordingly
```

Add structure loss.

```{r}
tmp1 <- mtbs.co %>% group_by(MTBS_YEAR) %>%
  summarize(burned_area = sum(MTBS_ACRES)) %>%
  ungroup() %>%
  add_row(MTBS_YEAR=2020, burned_area=678721.3) %>%
  add_row(MTBS_YEAR=2021, burned_area=32860) %>%
  rename(START_YEAR = MTBS_YEAR)
tmp2 <- ics.co %>% as_tibble() %>%
  group_by(START_YEAR) %>%
  summarize(destr_tot = sum(STR_DESTROYED_TOTAL),
            destr_res = sum(STR_DESTROYED_RES_TOTAL)) %>%
  ungroup() %>%
  add_row(START_YEAR = 2021, destr_tot = 991, destr_res = 969)
```


```{r warning=F, message=F, fig.width=5.25, fig.height=3}
f7 <- tmp1 %>%
  mutate(burned_km2 = burned_area*0.00404686) %>%
  ggplot(aes(x=START_YEAR, y=log(burned_area))) +
  geom_line() +
  geom_point(shape=21, size = 3, color = "gray30", fill = "red") +
  scale_x_continuous(limits=c(1984, 2021)) +
  geom_bar(data=tmp2, aes(x=START_YEAR, y=destr_res), stat="identity") +
  labs(x="\nIgnition Year", y="Burned Area (acres)\n", title="Colorado Burned Area (1984-2021)\n") +
  theme_minimal() +
  theme(plot.title = element_text(size = 14))
f7
ggsave(f7, file = paste("C:/Users/mccoo/OneDrive/mcook/ics209/figs/", "CO_BurnedArea_1984to2021_wRES.png", sep=""), 
       width=5, height=3, dpi = 700) # adjust dpi accordingly
```

