---
title: "ics+fired"
output: html_document
---

# ICS + FIRED: Linking incident command reports with satellite-derived fire events

```{r setup, include=F, echo=F, warning=F, error=F}
rm(list = ls())
knitr::opts_chunk$set(echo = TRUE)
source("setup.R")
```

First, we create a spatial SF points object from the Point of Origin (POO) coordinates.

```{r warning=F}

# incident.pts <- incidents %>%
#  rename(GID='...1') %>%
#  mutate(POO_LONGITUDE = if_else(is.na(POO_LONGITUDE), LRGST_FOD_LONGITUDE, POO_LONGITUDE),
#         POO_LATITUDE = if_else(is.na(POO_LATITUDE), LRGST_FOD_LATITUDE, POO_LATITUDE)) %>%
#  # remove rows with missing coordinates
#  filter(!is.na(POO_LONGITUDE)) %>%
#  # convert to spatial object
#  st_as_sf(coords=c("POO_LONGITUDE", "POO_LATITUDE"),crs=wgs.prj) %>%
#  # retain coordinates
#  mutate(POO_LONGITUDE = unlist(map(.$geometry,1)),
#         POO_LATITUDE = unlist(map(.$geometry,2))) %>%
#  # ensure transformation to WGS
#  st_set_crs(st_crs(wgs.prj)) %>%
#  # perform an intersection with our states (removing invalid points)
#  st_intersection(., states%>%dplyr::select(STUSPS)) %>%
#  # cast to point
#  st_cast("POINT") %>%
#  st_transform(st_crs(lambert.prj))
# # Write
# st_write(incident.pts,
#          "../../data/spatial/raw/wf-incidents/ics-209-plus-2.0/ics209plus-wf_incidents_spatial_us_1999to2020.gpkg",
#          driver="GPKG",delete_dsn=T)

```

```{r warning=F}

# Load the spatial points
incident.pts <- st_read("../../data/spatial/raw/wf-incidents/ics-209-plus-2.0/ics209plus-wf_incidents_spatial_us_1999to2020.gpkg") %>%
 # Filter 2001-2020 to match FIRED
 filter(START_YEAR>=2001) %>%
 # Remove some known "problem fires"
 filter(INCIDENT_ID != "2017_7211255_REDWOOD VALLEY INCIDENT",
        INCIDENT_ID != "2017_7293073_REDWOOD VALLEY INCIDENT") %>%
 # Make sure there are no duplicate records
 distinct(., INCIDENT_ID, .keep_all=TRUE)

# plot an example (Colorado)
ggplot() +
 geom_sf(data=incident.pts %>% filter(STUSPS=="CO") %>% arrange(FINAL_ACRES), 
         aes(color=FINAL_ACRES, size=FINAL_ACRES)) +
 viridis::scale_color_viridis(option="inferno",trans="log10") +
 scale_size(range = c(0.5,8)) +
 geom_sf(data=states %>% filter(STUSPS=="CO"), fill=NA) +
 coord_sf(crs=lambert.prj) +
 theme_minimal() +
 labs(color="Final Acres") +
 guides(color = guide_colourbar(title.position="top", barwidth = 10, barheight = 0.8, ticks=F,
                                label.theme = element_text(angle = 0, size = 8)),
        size="none") +
 theme(legend.position = "top")

```

## Single-fire events co-occurring with MTBS and FIRED perimeters (high confidence joins)

Now, we create a subset of wildfires that were not managed as complex incidents. We can do this by cross-checking against the complex associations table. We can also remove any rows where the FPA FOD indicates multiple fires, regardless of whether they were managed as complexes in ICS209s (keeping NULL values).

Join these records to MTBS perimeters, where possible. Keep only unique rows (i.e., remove rows where the MTBS perimeter is duplicated) and where the difference in recorded burn acres is less than 30,000. This ensure we are retaining only "high-confidence" joins. 

```{r}

# Load the complex associations tables
complex <- read_csv("data/tabular/raw/wf-incidents/ics-209-plus-2.0/ics209-plus-wf_complex_associations_1999to2020.csv")
cmplx14 <- read_csv("data/tabular/raw/wf-incidents/ics-209-plus-2.0/cpx-assocs-2014-2020.csv")

# Isolate incidents which were not managed as complexes
no.cmplx <- incident.pts %>% 
 # Filter incidents with 0 acres
 filter(FINAL_ACRES!=0,
        # remove incidents found in the complex associations table
        !INCIDENT_ID %in% complex$MEMBER_INCIDENT_ID,
        !INCIDENT_ID %in% cmplx14$MEMBER_INCIDENT_ID,
        # filter to fires where FPA-FOD shows 1 event or NULL
        FOD_FIRE_NUM==1 | is.na(FOD_FIRE_NUM))

# print the summary to make sure it worked
summary(no.cmplx$FOD_FIRE_NUM)
# check how many have MTBS ID
dim(no.cmplx%>%filter(!is.na(LRGST_MTBS_FIRE_INFO)))[1]

# clean the MTBS field
no.cmplx <- no.cmplx %>%
  mutate(Event_ID = gsub("\\s*\\([^\\)]+\\)","",
                         LRGST_MTBS_FIRE_INFO))

```

Now join the ICS209s to their MTBS perimeter. Filter by size/date to remove obvious issues.

```{r}
# Make sure CRS matches
mtbs <- mtbs %>% st_transform(st_crs(lambert.prj))
no.cmplx <- no.cmplx %>% st_transform(st_crs(lambert.prj))

# Join by "Event_ID"
mtbs.inci <- inner_join(
 no.cmplx %>% as_tibble() %>% dplyr::select(-geom),
 mtbs %>% as_tibble(), by="Event_ID"
) %>%
 # Handle duplicate MTBS perims
 distinct(Event_ID, .keep_all=T) %>%
 mutate(MTBS_ACRE_DIFF = abs(FINAL_ACRES - BurnBndAc),
        MTBS_PERC_DIFF = abs((FINAL_ACRES - BurnBndAc)/BurnBndAc) * 100,
        MTBS_DATE_DIFF = (abs(as.numeric(difftime(DISCOVERY_DATE, MTBS_Ig_Date, units = "days"))))) %>%
 # Remove joins if the acre difference is > 50,000 and ignition date difference is greater 31 days
 # Not exhaustive but handles some of the most glaring issues
 filter(MTBS_DATE_DIFF < 32,
        MTBS_ACRE_DIFF < 50000 | MTBS_PERC_DIFF < 50) %>%
 # Now cast back to multipolygon
 st_as_sf() %>%
 st_transform(lambert.prj) %>%
 st_make_valid() %>%
 st_cast("MULTIPOLYGON")

# Check the histogram of acre difference
summary(mtbs.inci$MTBS_ACRE_DIFF)
ggplot(data=mtbs.inci, aes(x=MTBS_ACRE_DIFF)) + 
 geom_histogram(bins=10,binwidth=1000)

# write this file out
st_write(mtbs.inci, "../../data/spatial/mod/ics-fired/mtbs_perims_w_incidents.gpkg", delete_dsn=T)

```

Now find co-occurring FIRED events using MTBS perimeters. Do this for events with high-confidence MTBS join (i.e., same incident name and no duplicate records). Also, use the largest overlap and perform the join by fire year to try and find the most likely event.

```{r warning=F}

# Set the geometries
st_agr(events) <- "constant"
st_agr(mtbs.inci) <- "constant"

# Set CRS and cast to be explicit
events <- events %>%
 st_transform(st_crs(lambert.prj)) %>%
 st_make_valid() %>%
 st_cast("MULTIPOLYGON")

# Loop through years, perform spatial join w/ largest overlap
join = c()
years = seq(2001,2020,1)
for (y in 1:length(years)){
 print(years[y])
 a <- events[events$ig_year == years[y],]
 st_agr(a) = "constant"
 b <- mtbs.inci[mtbs.inci$START_YEAR == years[y],]
 st_agr(b) = "constant"
 join[[y]] <- st_join(b,a,join=st_intersects,largest=TRUE,left=FALSE)
}

# Merge the list of lists
df <- bind_rows(join) %>%
 # remove rows without a join
 filter(!is.na(id)) %>%
 mutate(tot_ar_ac = tot_ar_km2*247.105,
        FIRED_ACRE_DIFF = abs(BurnBndAc-tot_ar_ac),
        FIRED_DATE_DIFF = (abs(as.numeric(difftime(DISCOVERY_DATE, ig_date, units = "days")))),
        FIRED_PERC_DIFF = abs((tot_ar_ac - BurnBndAc)/BurnBndAc) * 100) %>%
 # Create simple data frame
 as_tibble() %>%
 dplyr::select(-geom)

# Any duplicate FIRED IDs?
if(dim(df %>% group_by(id) %>% filter(n()>1))[1] > 1 |
   dim(df %>% group_by(INCIDENT_ID) %>% filter(n()>1))[1] > 1){
 print("Duplicate IDs found on first pass")
} else (print("No duplicate IDs on first pass"))

########################################

# Check the histogram of acre difference
summary(df$FIRED_ACRE_DIFF)
ggplot(data=df, aes(x=FIRED_ACRE_DIFF)) + 
 geom_histogram(bins=10,binwidth=10000)

# Check the histogram of date difference
summary(df$FIRED_DATE_DIFF)
ggplot(data=df, aes(x=FIRED_DATE_DIFF))+
 geom_histogram(bins=10,binwidth=10)

########################################

# Filter out the most glaring issues 
# (acre difference > 50,000, date difference > 25 days)
# 96% of records are within 25 days
# Keep only distinct matches
df <- df %>% 
 filter(FIRED_DATE_DIFF <= 25,
        FIRED_ACRE_DIFF < 50000 | FIRED_PERC_DIFF < 50) %>%
 group_by(id) %>%
 # If there are duplicates, take the nearest in size
 slice(which.min(FIRED_PERC_DIFF)) %>%
 ungroup() %>%
 # Keep distinct
 distinct(., id, .keep_all = TRUE)

# Any duplicate FIRED IDs?
if(dim(df %>% group_by(id) %>% filter(n()>1))[1] > 1 |
   dim(df %>% group_by(INCIDENT_ID) %>% filter(n()>1))[1] > 1){
 print("Duplicate IDs found on second pass")
} else (print("No duplicate IDs on second pass"))

# Grab a summary of the difference fields
summary(df$FIRED_ACRE_DIFF)
summary(df$FIRED_DATE_DIFF)

# Now, Isolate the FIRED events, create cleaned table,export
# Add in the FIRED_ID and other fields to the ICS-209s
events.ics <- events %>% 
 as_tibble() %>%
 dplyr::select(id,geom) %>%
 filter(id %in% df$id) %>%
 inner_join(.,df,by="id") %>%
 rename(FIRED_ID = id) %>%
 # Now cast back to multipolygon
 st_as_sf() %>%
 st_transform(lambert.prj) %>%
 st_make_valid() %>%
 st_cast("MULTIPOLYGON")

dim(events.ics %>% group_by(FIRED_ID) %>% filter(n()>1))[1]
dim(events.ics %>% group_by(INCIDENT_ID) %>% filter(n()>1))[1]

# Need to double check column names for exporting
if(any(duplicated(toupper(names(events.ics))))){
 print(which(duplicated(toupper(names(events.ics)))))
} else {
 print("Column names OK")
}

# Write to file
st_write(events.ics,"../../data/spatial/mod/ics-fired/ics209plus_fired_by_mtbs.gpkg",delete_dsn=TRUE)

# Tidy up
rm(mtbs,df,a,b,join)
# abs((264155 - 348772)/348772) * 100
```

Some "validation". Run an LM and check on r2 between ICS and FIRED acres.

```{r}

# Create a LM to show the relationship between ICS and FIRED acres
lm.fit <- lm(FINAL_ACRES ~ tot_ar_ac, data=events.ics)
r2 = round(summary(lm.fit)$r.squared, 2)
summary(lm.fit)
print("~~~~~~~~~")

# Predict
pred <- predict(lm.fit, interval="confidence")
pred <- cbind(events.ics,pred)
ggplot(data=pred, aes(x=tot_ar_ac,y=FINAL_ACRES)) +
 geom_point() +
 geom_ribbon(aes(ymin=lwr,ymax=upr),alpha=0.15)+
 geom_line(aes(y=fit),size=1)+
 labs(title="Relationship b/w ICS and FIRED Acres \n(MTBS-based Join)",
      subtitle=paste0("R-sq: ",r2),
      x="FIRED Acres",y="ICS Acres")
rm(r2,lm.fit)

```

Look at the average distance between incident POO and FIRED centroid.

```{r}

library(nngeo, quietly=T)

centroid <- events %>%
 filter(id %in% events.ics$FIRED_ID) %>% 
 st_centroid() %>%
 st_transform(st_crs(lambert.prj))

pts <- incident.pts%>%filter(INCIDENT_ID %in% events.ics$INCIDENT_ID)

nn <- st_nn(pts, centroid, k=1, returnDist = T)

nn.df <- as.matrix(nn$dist)
head(nn.df)
nn.df <- as.data.frame(nn.df)%>%
 mutate(V1 = as.numeric(V1))
print("~~~~~~")
paste0("Mean Distance: ", round(mean(nn.df$V1),2)," meters")
paste0("Standard Deviation Distance: ", round(sd(nn.df$V1),2))

rm(centroid,pts,nn,nn.df)

```

## Linking incident command report to FIRED using buffer around POO coordinates and spatial/temporal filters.

The average distance between the incident POO and the centroid of the FIRED event in our first round is 3.6km with high variability (stdev = 4.6km). This is a decent sample because the joins are high-confidence.

Work with complex fires first: Starting with a 20km buffer around incident POO, identify all FIRED events within the buffer and then slice the data frame to find rows where the ignition dates are within 25 days.

First, do some cleaning. Isolate complex fires, handle manual edits.

```{r}

# Isolate complex fires
# Handle some of the known "duplicates"

# Read in the updated complex assoc. table (2014-2020)
# Retrieve only the complex INCIDENT_ID from the spatial points
cmplx14 <- cmplx14 %>%
 # Remove "problem" fires if known
 filter(MEMBER_INCIDENT_ID != "2017_7211255_REDWOOD VALLEY INCIDENT")

# Merge 1999-2013 and 2014-2020 complex records
cmplx1 <- incident.pts %>%
 filter(INCIDENT_ID %in% cmplx14$CPLX_INCIDENT_ID)
cmplx2 <- incident.pts %>%
 filter(START_YEAR<2014, 
        INCIDENT_ID %in% complex$CPLX_INCIDENT_ID)

# Bind them together
cmplx3 <- bind_rows(cmplx1,cmplx2)

# Also add in fires where FOD_FIRE_NUM > 1
cmplx4 <- incident.pts %>%
 filter(!INCIDENT_ID %in% cmplx3$INCIDENT_ID,
        FOD_FIRE_NUM > 1)

# Final bind
cmplx <- bind_rows(cmplx3,cmplx4)

# Create the member ID table
cmplx.member <- incident.pts %>%
 filter(
  INCIDENT_ID %in% cmplx14$MEMBER_INCIDENT_ID)
temp <- incident.pts %>%
 filter(
  START_YEAR < 2014,
  INCIDENT_ID %in% complex$MEMBER_INCIDENT_ID)
# Bind together
cmplx.member <- bind_rows(cmplx.member,temp)

 # Tidy up
rm(cmplx1,cmplx2,cmplx3,cmplx4,temp)

```

Now buffer the complex incident summary reports by 20km (accounting for mean and standard deviation distance in MTBS-based join). Filter by ignition date and size differences.

```{r}

# Filter FIRED events
# Remove events which have been joined already
st_agr(events)="constant"
events_ <- events %>%
 mutate(tot_ar_ac = tot_ar_km2*247.105) %>%
 filter(!id %in% events.ics$FIRED_ID) %>%
 st_transform(st_crs(lambert.prj))

# Buffer complex incident points by 20km
buffer <- cmplx %>% 
 filter(!INCIDENT_ID %in% events.ics$INCIDENT_ID) %>%
 st_buffer(20000) %>% st_transform(st_crs(lambert.prj))

# Loop through years
join2 = c()
years = seq(2001,2020,1)
for (y in 1:length(years)){
 print(years[y])
 a <- events_[events_$ig_year == years[y],]
 st_agr(a) = "constant"
 b <-buffer[buffer$START_YEAR == years[y],]
 st_agr(b) = "constant"
 join2[[y]] <- st_intersection(b,a)
}

# Bind the results
df <- bind_rows(join2) %>%
 # remove rows without a join
 filter(!is.na(id)) %>%
 mutate(tot_ar_ac = tot_ar_km2*247.105,
        FIRED_ACRE_DIFF = abs(FINAL_ACRES-tot_ar_ac),
        FIRED_PERC_DIFF = abs((tot_ar_ac - FINAL_ACRES)/FINAL_ACRES) * 100,
        FIRED_DATE_DIFF = (abs(as.numeric(difftime(DISCOVERY_DATE, ig_date, units = "days"))))) %>%
 # Create simple data frame
 st_set_geometry(NULL) %>% 
 as_tibble()

head(df)

# Get a summary of the date and acre differences
summary(df$FIRED_DATE_DIFF)
summary(df$FIRED_ACRE_DIFF)

# Now filter the data frame
df <- df %>%
 # Keep near-in-time
 filter(FIRED_DATE_DIFF <= 25,
        FIRED_ACRE_DIFF < 50000 | FIRED_PERC_DIFF < 50) %>%
 # Group by INCIDENT_ID
 group_by(INCIDENT_ID) %>%
 # Keep the closest in size
 slice(which.min(FIRED_PERC_DIFF)) %>%
 ungroup() %>% group_by(id) %>%
 slice(which.min(FIRED_PERC_DIFF)) %>% 
 ungroup()

head(df)

# Checks
dim(df %>% group_by(INCIDENT_ID) %>% filter(n()>1))[1]
dim(df %>% group_by(id) %>% filter(n()>1))[1]

########################################################

# Check the histogram of acre difference
summary(df$FIRED_ACRE_DIFF)
ggplot(data=df, aes(x=FIRED_ACRE_DIFF)) + 
 geom_histogram(bins=10,binwidth=10000)

# Check the histogram of acre % difference
summary(df$FIRED_PERC_DIFF)
ggplot(data=df, aes(x=FIRED_PERC_DIFF)) + 
 geom_histogram(bins=10,binwidth=10)

# Check the histogram of date difference
summary(df$FIRED_DATE_DIFF)
ggplot(data=df, aes(x=FIRED_DATE_DIFF))+
 geom_histogram(bins=10,binwidth=10)

########################################################

if(
 dim(df %>% group_by(INCIDENT_ID) %>% filter(n()>1))[1] > 1 |
 dim(df %>% group_by(id) %>% filter(n()>1))[1] > 1
) {
 # Filter using conservative thresholds
 # This should remove glaring issues
 # within 5k acres (useful for smaller fires),
 # or, 30% difference (useful for larger fires)
 df <- df %>%
  filter(FIRED_ACRE_DIFF < 5000 | FIRED_PERC_DIFF < 35)
} else (print("No duplicate fire joins detected ..."))

df <- df %>%
  filter(FIRED_ACRE_DIFF < 5000 | FIRED_PERC_DIFF < 50)

# Get the join as FIRED polygons
events.ics2 <- events %>%
 as_tibble() %>%
 dplyr::select(id,geom) %>%
 filter(id %in% df$id) %>%
 inner_join(.,df,by="id") %>%
 rename(FIRED_ID = id) %>%
 # Now cast back to multipolygon
 st_as_sf() %>%
 st_transform(lambert.prj) %>%
 st_make_valid() %>%
 st_cast("MULTIPOLYGON")

# Write to file
st_write(events.ics2,"../../data/spatial/mod/ics-fired/ics209plus_fired_by_buffer_cmplx.gpkg",delete_dsn=TRUE)

# Now clean up the complex table
# Remove MEMBER_INCIDENT_IDs that are related to any complex fires that made it through the join
cmplx.rem <- cmplx14 %>%
 filter(!CPLX_INCIDENT_ID %in% events.ics2$INCIDENT_ID)
temp <- complex %>%
 filter(!CPLX_INCIDENT_ID %in% cmplx.rem$CPLX_INCIDENT_ID,
        !CPLX_INCIDENT_ID %in% events.ics2$INCIDENT_ID)
cmplx.rem <- bind_rows(cmplx.rem,temp)

# Tidy up
rm(a,b,buffer,df,join2,temp)

```

Validation.

```{r}

# Create a LM to show the relationship between ICS and FIRED acres
lm.fit <- lm(FINAL_ACRES ~ tot_ar_ac, data=events.ics2)
r2 = round(summary(lm.fit)$r.squared, 2)
summary(lm.fit)

# Predict
events.ics2$pred2 <- predict(lm.fit)
ggplot(data=events.ics2, aes(x=tot_ar_ac,y=FINAL_ACRES)) +
 geom_point() +
 geom_line(aes(y=pred2),size=1)+
 labs(title="Relationship b/w ICS and FIRED Acres \n(Buffer-based Join, Complexes Only)",
      subtitle=paste0("R-sq: ",r2),
      x="FIRED Acres",y="ICS Acres")
rm(r2,lm.fit)

```

Now work with the non-complex incidents which have not yet been joined through the MTBS link. Apply the 20km buffer to these incidents and identify overlapping FIRED events.

```{r}

# Remove incidents that were joined in Step 1&2. 
incis_no.cmplx <- no.cmplx %>%
 filter(!INCIDENT_ID %in% events.ics$INCIDENT_ID,
        !INCIDENT_ID %in% events.ics2$INCIDENT_ID) %>%
 st_transform(st_crs(lambert.prj))

# Remove joined events
events_ <- events_ %>% filter(!id %in% events.ics2$FIRED_ID)

# Buffer incident points by 20km
buffer <- incis_no.cmplx %>% st_buffer(20000) %>% st_transform(st_crs(lambert.prj))

# Loop through years
join = c()
years = seq(2001,2020,1)
for (y in 1:length(years)){
 print(years[y])
 a <- events_[events_$ig_year == years[y],]
 st_agr(a) = "constant"
 b <-buffer[buffer$START_YEAR == years[y],]
 st_agr(b) = "constant"
 join[[y]] <- st_intersection(b,a)
}

# Bind the results
df <- bind_rows(join) %>%
 st_set_geometry(NULL) %>% 
 as_tibble() %>%
 # Create the difference attributes
 mutate(FIRED_ACRE_DIFF = abs(FINAL_ACRES-tot_ar_ac),
        FIRED_PERC_DIFF = abs((tot_ar_ac - FINAL_ACRES)/FINAL_ACRES) * 100,
        FIRED_DATE_DIFF = (abs(as.numeric(difftime(DISCOVERY_DATE, ig_date, units = "days")))))

head(df)

# Get a summary of the date and acre differences
summary(df$FIRED_DATE_DIFF)
summary(df$FIRED_ACRE_DIFF)

# Filter using differences in ignition date and final size
# Group by incident and compare
df <- df %>%
 # Keep near-in-time
 filter(FIRED_DATE_DIFF <= 25,
        FIRED_ACRE_DIFF < 50000 | FIRED_PERC_DIFF < 50) %>%
 group_by(INCIDENT_ID) %>%
 # Keep the closest in size
 slice(which.min(FIRED_PERC_DIFF)) %>%
 ungroup() %>% group_by(id) %>%
 slice(which.min(FIRED_PERC_DIFF)) %>% 
 ungroup()

head(df)

# Checks
dim(df %>% group_by(INCIDENT_ID) %>% filter(n()>1))[1]
dim(df %>% group_by(id) %>% filter(n()>1))[1]

########################################

# Check the histogram of acre difference
summary(df$FIRED_ACRE_DIFF)
ggplot(data=df, aes(x=FIRED_ACRE_DIFF)) + 
 geom_histogram(bins=10,binwidth=10000)

# Check the histogram of date difference
summary(df$FIRED_DATE_DIFF)
ggplot(data=df, aes(x=FIRED_DATE_DIFF))+
 geom_histogram(bins=10,binwidth=10)

########################################

# Filter using conservative thresholds
# This should remove glaring issues
# within 5k acres (useful for smaller fires),
# Smaller fires may have less than a 50k acre difference but the % difference is massive
# or, 30% difference (useful for larger fires)
df <- df %>%
 filter(FIRED_ACRE_DIFF < 5000 | FIRED_PERC_DIFF < 50)

# Get the join as FIRED polygons
events.ics3 <- events %>%
 as_tibble() %>%
 dplyr::select(id,geom) %>%
 filter(id %in% df$id) %>%
 inner_join(.,df,by="id") %>%
 rename(FIRED_ID = id) %>%
 # Now cast back to multipolygon
 st_as_sf() %>%
 st_transform(lambert.prj) %>%
 st_make_valid() %>%
 st_cast("MULTIPOLYGON")

# Write to file
st_write(events.ics3,"../../data/spatial/mod/ics-fired/ics209plus_fired_by_buffer_no_cmplx.gpkg",delete_dsn=TRUE)

# Tidy
rm(a,b,df,join,buffer,incis_no.cmplx)
```

"Validation".

```{r}

# Create a LM to show the relationship between ICS and FIRED acres
lm.fit <- lm(FINAL_ACRES ~ tot_ar_ac, data=events.ics3)
r2 = round(summary(lm.fit)$r.squared, 2)
summary(lm.fit)

# Predict
events.ics3$pred3 <- predict(lm.fit)
ggplot(data=events.ics3, aes(x=tot_ar_ac,y=FINAL_ACRES)) +
 geom_point() +
 geom_line(aes(y=pred3),size=1)+
 labs(title="Relationship b/w ICS and FIRED Acres \n(Buffer-based Join, Non-Complexes)",
      subtitle=paste0("R-sq: ",r2),
      x="FIRED Acres",y="ICS Acres")
rm(r2,lm.fit)

```

## Linking remaining complex incidents to FIRED

First, create the master data frame to this point (bind rows on the first three joins). Check what % is joined for a couple key stats.

```{r}

events.ics.master <- bind_rows(events.ics,events.ics2,events.ics3)

# Create a temp file to remove incidents which are part of already joined complexes
cmplx1 <- cmplx14 %>% 
 filter(CPLX_INCIDENT_ID %in% events.ics.master$INCIDENT_ID)

cmplx2 <- complex %>% filter(!CPLX_INCIDENT_ID %in% cmplx1$CPLX_INCIDENT_ID,
                             CPLX_INCIDENT_ID %in% events.ics.master$INCIDENT_ID)
dim(cmplx2%>%filter(CPLX_INCIDENT_ID %in% cmplx1$CPLX_INCIDENT_ID))
dim(cmplx2%>%filter(MEMBER_INCIDENT_ID %in% cmplx1$MEMBER_INCIDENT_ID))

cmplx.joined <- bind_rows(cmplx1,cmplx2)

rm(cmplx1,cmplx2)

dim(incident.pts%>%filter(INCIDENT_ID %in% cmplx.joined$CPLX_INCIDENT_ID))

# Create the df with complexes and member fires removed
incident.pts_ <- incident.pts %>%
 filter(!INCIDENT_ID %in% cmplx.joined$MEMBER_INCIDENT_ID, 
         !INCIDENT_NAME %in% cmplx.joined$FIRE_NAME)
temp <- incident.pts %>%
 filter(INCIDENT_ID %in% cmplx.joined$CPLX_INCIDENT_ID |
         INCIDENT_NAME %in% cmplx.joined$ICS_COMPLEX_NAME)
incident.pts_ <- bind_rows(incident.pts_,temp) %>%
 distinct(., INCIDENT_ID,.keep_all=TRUE)
rm(temp)

# Checks
dim(events.ics.master %>% group_by(INCIDENT_ID) %>% filter(n()>1))[1]
dim(incident.pts_ %>% group_by(INCIDENT_ID) %>% filter(n()>1))[1]

# Get some stats
sum(events.ics.master$STR_DESTROYED_RES_TOTAL) / sum(incident.pts_$STR_DESTROYED_RES_TOTAL) * 100
sum(events.ics.master$FINAL_ACRES) / sum(incident.pts_$FINAL_ACRES) * 100
(sum(events.ics.master$PROJECTED_FINAL_IM_COST,na.rm=TRUE)/1e9) / 
 (sum(incident.pts_$PROJECTED_FINAL_IM_COST,na.rm=TRUE)/1e9) * 100

rm(events.ics, events.ics2, events.ics3)
```

Now we want to work with remaining incidents. Importantly, we are now working with the incident summary table with complex member fires removed as these should be accounted for in the successful complex fire joins (ex/ August Complex, Central LNU, Southern LNU, etc.).

In this workflow, we will use KNN to identify possible matching fire events. The join returns all FIRED centroid within a 50km distance of the incident POO (by year). 

```{r}

# Filter joined FIRED events
events_ <- events_ %>% filter(!id %in% events.ics.master$FIRED_ID)

# Grab remaining incidents
incis_ <- incident.pts_ %>%
 filter(!INCIDENT_ID %in% events.ics.master$INCIDENT_ID)

# Grab the centroid
centroid <- events_ %>% st_centroid()

# Loop through years
join = c()
years = seq(2001,2020,1)
for (y in 1:length(years)){
 print(years[y])
 a <- centroid[centroid$ig_year == years[y],]
 st_agr(a) = "constant"
 b <- incis_[incis_$START_YEAR == years[y],]
 st_agr(b) = "constant"
 join[[y]] <- st_join(b,a,st_nn,k=50,maxdist=50000,left=TRUE)
}

# Bind the results
df <- bind_rows(join) %>%
 st_set_geometry(NULL) %>% as_tibble() %>%
 # Create the difference attributes
 mutate(FIRED_ACRE_DIFF = abs(FINAL_ACRES-tot_ar_ac),
        FIRED_PERC_DIFF = abs((tot_ar_ac - FINAL_ACRES)/FINAL_ACRES) * 100,
        FIRED_DATE_DIFF = (abs(as.numeric(difftime(DISCOVERY_DATE, ig_date, units = "days")))))
head(df)

# Filter using differences in ignition date and final size
# Group by incident and compare
df <- df %>%
 # Keep near-in-time
 filter(FIRED_DATE_DIFF <= 25,
        FIRED_ACRE_DIFF < 50000 | FIRED_PERC_DIFF < 55) %>% # handles the East Troublesome Fire
 group_by(INCIDENT_ID) %>%
 # Keep the closest in size
 slice(which.min(FIRED_PERC_DIFF)) %>%
 ungroup() %>% group_by(id) %>%
 slice(which.min(FIRED_PERC_DIFF)) %>% 
 ungroup()
head(df)

# Checks
dim(df %>% group_by(INCIDENT_ID) %>% filter(n()>1))[1]
dim(df %>% group_by(id) %>% filter(n()>1))[1]

########################################
# Check the histogram of acre difference
summary(df$FIRED_ACRE_DIFF)
ggplot(data=df, aes(x=FIRED_ACRE_DIFF)) + 
 geom_histogram(bins=10,binwidth=10000)

# Check the histogram of date difference
summary(df$FIRED_DATE_DIFF)
ggplot(data=df, aes(x=FIRED_DATE_DIFF))+
 geom_histogram(bins=10,binwidth=10)
########################################

# Now filter thresholds
# This should remove glaring issues
df <- df %>%
 filter(FIRED_ACRE_DIFF < 5000 | FIRED_PERC_DIFF < 55)

# Get the join as FIRED polygons
events.ics4 <- events %>%
 as_tibble() %>%
 dplyr::select(id,geom) %>%
 filter(id %in% df$id) %>%
 inner_join(.,df,by="id") %>%
 rename(FIRED_ID = id) %>%
 # Now cast back to multipolygon
 st_as_sf() %>%
 st_transform(lambert.prj) %>%
 st_make_valid() %>%
 st_cast("MULTIPOLYGON")

# Write to file
st_write(events.ics4,"../../data/spatial/mod/ics-fired/ics209plus_fired_by_buffer_remaining.gpkg",delete_dsn=TRUE)

# Tidy
rm(join,df,a,b)
```

Validation.

```{r}

# Create a LM to show the relationship between ICS and FIRED acres
lm.fit <- lm(FINAL_ACRES ~ tot_ar_ac, data=events.ics4)
r2 = round(summary(lm.fit)$r.squared, 2)
summary(lm.fit)

# Predict
events.ics4$pred4 <- predict(lm.fit)
ggplot(data=events.ics4, aes(x=tot_ar_ac,y=FINAL_ACRES)) +
 geom_point() +
 geom_line(aes(y=pred4),size=1)+
 labs(title="Relationship b/w ICS and FIRED Acres \n(Buffer-based Join, Remaining)",
      subtitle=paste0("R-sq: ",r2),
      x="FIRED Acres",y="ICS Acres")
rm(r2,lm.fit)

```

Merge. This represents what is likely the most accurate subset of incidents/FIRED events. 

Grab some statistics.

```{r}

events.ics.master_ <- bind_rows(events.ics.master,events.ics4)

# Check for duplicate records ...
dim(events.ics.master_ %>% group_by(INCIDENT_ID) %>% filter(n()>1))[1]
dim(events.ics.master_ %>% group_by(FIRED_ID) %>% filter(n()>1))[1]

# Handle duplicate records
events.ics.master_ <- events.ics.master_ %>% distinct(., INCIDENT_ID, .keep_all=TRUE)

# Get some stats
# Make sure to remove member incidents from complex that have been joined
sum(events.ics.master_$STR_DESTROYED_RES_TOTAL) / sum(incident.pts_$STR_DESTROYED_RES_TOTAL) * 100
sum(events.ics.master_$FINAL_ACRES) / sum(incident.pts_$FINAL_ACRES) * 100
(sum(events.ics.master_$PROJECTED_FINAL_IM_COST,na.rm=TRUE)/1e9) / 
 (sum(incident.pts_$PROJECTED_FINAL_IM_COST,na.rm=TRUE)/1e9) * 100

rm(events.ics4)
```

Validation. Final.

```{r}

# Create a LM to show the relationship between ICS and FIRED acres
lm.fit <- lm(FINAL_ACRES ~ tot_ar_ac, data=events.ics.master_)
r2 = round(summary(lm.fit)$r.squared, 2)
summary(lm.fit)

# Predict
events.ics.master_$pred5 <- predict(lm.fit)
ggplot(data=events.ics.master_, aes(x=tot_ar_ac,y=FINAL_ACRES)) +
 geom_point() +
 geom_line(aes(y=pred5),size=1)+
 labs(title="Relationship b/w ICS and FIRED Acres \n(Final Join)",
      subtitle=paste0("R-sq: ",r2),
      x="FIRED Acres",y="ICS Acres")
rm(r2,lm.fit)

```

Add in the FIRED point of origin coordinates. Final bind and export.

```{r}

# Create the master table
events.ics.master_ <- events.ics.master_ %>%
 dplyr::select(-c(pred2,pred3,pred4)) %>%
 rename(PREDICTED_ACRES = pred5)

# Add in FIRED POO
ig.coords <- st_read("../../../FIRED/data/spatial/raw/conus-ak_to2022_ig_points_wgs.gpkg")

events.ics.master_ <- events.ics.master_ %>%
 left_join(ig.coords, by="FIRED_ID")

# Write out the master events data so far
st_write(events.ics.master_, "../../data/spatial/mod/ics-fired/ics209plus_fired_combined.gpkg",
         delete_dsn=TRUE)

```

