---
title: "ics+fired"
output: html_document
---

# Linking ICS-209-PLUS to FIRED (ICS+FIRED)

```{r setup, include=F, echo=F, warning=F, error=F}
# rm(list = ls())
knitr::opts_chunk$set(echo = TRUE)
source("setup.R")
```

```{r}
# Grab a glimpse of the data
glimpse(ics.conus)
```

In CONUS, there were 61,097 homes destroyed across 2,360 wildfire incidents representing 7.19% of total wildfire incidents during this time period (1999-2020). In the western U.S., 52,119 homes were lost to wildfire in just 1,007 individual incidents. This represents only 6.7% of west-wide incidents. Wildfire in the west accounts for 85.3% of home loss (1999-2020) across just 3.1% of wildfire incidents.

```{r}
paste("Num. of homes destroyed by wildfire (CONUS): ", 
      as.integer(sum(ics.conus$STR_DESTROYED_RES_TOTAL)))
paste("Num. of destructive incidents (CONUS): ", 
      dim(ics.conus %>% filter(STR_DESTROYED_RES_TOTAL > 0))[1], sep="")
paste("Num. of homes destroyed by wildfire (WEST): ", sum(ics.west$STR_DESTROYED_RES_TOTAL))
paste("Num. of destructive incidents (WEST): ", 
      dim(ics.west %>% filter(STR_DESTROYED_RES_TOTAL > 0))[1], sep="")
paste("Percentage destroyed homes in the West: ",
      round(sum(ics.west$STR_DESTROYED_RES_TOTAL)/sum(ics.conus$STR_DESTROYED_RES_TOTAL) * 100, 2), sep="")
```

Plot destruction by state.

```{r fig.width=4, fig.height=2.5}
# Group summary by state
state.sum <- ics.conus %>% as_tibble() %>%
  filter(STR_DESTROYED_RES_TOTAL > 0) %>%
  group_by(STUSPS) %>%
  summarize(str_des_res = sum(STR_DESTROYED_RES_TOTAL)) %>%
  inner_join(states%>%as_tibble(), ., by="STUSPS") %>%
  st_as_sf() 
  # Plot
ggplot() +
  geom_sf(data=state.sum, aes(fill=str_des_res)) +
  scale_fill_viridis_c(option="magma", trans="log10") +
  guides(fill="none") +
  theme_void()
rm(state.sum)
```

Texas and Oklahoma account for a large percentage of home loss. Tennessee is also in the top 15 most destructive states. By including TX and OK in the analysis we can capture 93% of all wildfire-related home loss and 91% of total burned area from the incident command reports. 

```{r}
west.plus <- c("AZ", "CO", "NV", "WY", "CA", "ID", "WA", "OR", "NM", "MT", "UT", "TX", "OK")
ics.west.plus <- ics.conus %>% filter(STUSPS %in% west.plus)
## Subset FIRED as well
fired.west.plus <- fired.conus %>% filter(STUSPS %in% west.plus)
## Homes destroyed
print("Homes destroyed: ...")
sum(ics.west$STR_DESTROYED_RES_TOTAL) / sum(ics.conus$STR_DESTROYED_RES_TOTAL) * 100
sum(ics.west.plus$STR_DESTROYED_RES_TOTAL)
sum(ics.west.plus$STR_DESTROYED_RES_TOTAL) / sum(ics.conus$STR_DESTROYED_RES_TOTAL) * 100
## Burned area
print("Burned area (acres): ...")
sum(ics.west$FINAL_ACRES) / sum(ics.conus$FINAL_ACRES) * 100
sum(ics.west.plus$FINAL_ACRES)
sum(ics.west.plus$FINAL_ACRES) / sum(ics.conus$FINAL_ACRES) * 100
```

```{r}
paste("Num. of homes destroyed by wildfire (CONUS): ", 
      as.integer(sum(ics.conus$STR_DESTROYED_RES_TOTAL)))
paste("Num. of destructive incidents (CONUS): ", 
      dim(ics.conus %>% filter(STR_DESTROYED_RES_TOTAL > 0))[1], sep="")
paste("Num. of homes destroyed by wildfire (WEST): ", sum(ics.west.plus$STR_DESTROYED_RES_TOTAL))
paste("Num. of destructive incidents (WEST): ", 
      dim(ics.west.plus %>% filter(STR_DESTROYED_RES_TOTAL > 0))[1], sep="")
paste("Percentage destroyed homes in the West: ",
      round(sum(ics.west.plus$STR_DESTROYED_RES_TOTAL)/sum(ics.conus$STR_DESTROYED_RES_TOTAL) * 100, 2), sep="")
```

Map structures destroyed and burned area.

```{r fig.height=5, fig.width=5}
map1 <- ggplot() +
  geom_sf(data=ics.west.plus%>%filter(STR_DESTROYED_RES_TOTAL==0), color="gray60", size=0.3, alpha=0.9) +
  geom_sf(data=states%>%filter(STUSPS %in% states.sub), fill=NA, color="gray20") +
  geom_sf(data=ics.west.plus%>%filter(STR_DESTROYED_RES_TOTAL>0)%>%arrange(STR_DESTROYED_RES_TOTAL),
          aes(color=STR_DESTROYED_RES_TOTAL, size=FINAL_ACRES), alpha=0.8) +
  scale_size(range = c(1,6)) +
  scale_color_viridis_c(option="inferno", trans="log10") +
  labs(color = "Homes Destroyed") +
  coord_sf(crs = st_crs(ics.conus)) +
  theme_void() +
  guides(color = guide_colourbar(position="top", barwidth = 0.4, barheight = 6, ticks=F,
                                 label.position = "right", title.position = "left",
                                 label.theme = element_text(angle = 0, size = 8)),
         size="none") +
  theme(legend.title = element_text(angle = 90, size=10),
        legend.position=c(0.85, 0.7))
map1
ggsave(map1, file = paste("../../figs/Map1-ICS209_HomesDestroyed_Area_WestPlus.png", sep=""), 
       width=5, height=5, dpi = 350) # adjust dpi accordingly
```

Homes destroyed by incident.

```{r fig.width=4, fig.height=2.5, warning=F}
## reorder factors
ics.west.plus$INCIDENT_NAME <- 
  reorder(ics.west.plus$INCIDENT_NAME, ics.west.plus$STR_DESTROYED_RES_TOTAL)
ics.west.plus$INCIDENT_NAME <- 
  factor(ics.west.plus$INCIDENT_NAME, levels=rev(levels(ics.west.plus$INCIDENT_NAME)))
# Grab the top 15
top10 <- ics.west.plus %>% slice_max(STR_DESTROYED_RES_TOTAL, n=10)
# Plot and label
f1 <- ggplot(data=ics.west.plus, aes(x=START_YEAR, y=STR_DESTROYED_RES_TOTAL, 
                                     group=factor(INCIDENT_NAME), 
                                     fill=STR_DESTROYED_RES_TOTAL), color="grey90") +
  geom_bar(stat="identity", position='stack') +
  scale_fill_viridis_c(option="inferno", trans="log10") +
  labs(
    x="", y="Homes Destroyed\n", fill="") +
    # title="Homes destroyed by Western wildfires (1999-2020)\n",
    # caption="Maxwell C. Cook, PhD Student, Department of Geography,
    # University of Colorado, Boulder\nData source: St. Denis et al., 2020 (updated)") +
  # ggrepel::geom_text_repel(data=. %>% filter(INCIDENT_ID %in% top10$INCIDENT_ID), 
  #                          aes(label=INCIDENT_NAME), position=position_jitterdodge(0.5), size = 2) +
  theme_minimal() +
  theme(plot.title = element_text(size = 9),
        plot.subtitle = element_text(size=7),
        plot.caption = element_text(size=6),
        axis.title.x = element_text(size=8),
        axis.title.y = element_text(size=8),
        axis.text.y = element_text(size=7),
        axis.text.x = element_text(size=7),
        legend.position="none")
f1
ggsave(f1, file = "../../figs/Westwide-PLUS_HomeLoss_byIncident_1999to2020.png",
       width=4, height=2.5, dpi = 300) # adjust dpi accordingly
rm(top10)
```

Burned area (option 1).

```{r fig.width=4, fig.height=2.5, warning=F}
## reorder factors
ics.west.plus$INCIDENT_NAME <- 
  reorder(ics.west.plus$INCIDENT_NAME, ics.west.plus$FINAL_ACRES)
ics.west.plus$INCIDENT_NAME <- 
  factor(ics.west.plus$INCIDENT_NAME, levels=rev(levels(ics.west.plus$INCIDENT_NAME)))
## Plot it
f2 <- ggplot(data=ics.west.plus, aes(x=START_YEAR, y=FINAL_ACRES, 
                                     group=factor(INCIDENT_NAME), 
                                     fill=FINAL_ACRES), color="black") +
  geom_bar(stat='identity', position="stack") +
  scale_fill_viridis_c(option="rocket", trans="sqrt") +
  scale_y_continuous(labels = scales::label_number(suffix = " M", scale = 1e-6)) +
  theme_minimal() +
  labs(x="", y="Burned Area (acres)") +
  theme(plot.title = element_text(size = 9),
        plot.subtitle = element_text(size=7),
        plot.caption = element_text(size=6),
        axis.title.x = element_text(size=8),
        axis.title.y = element_text(size=8),
        axis.text.y = element_text(size=7),
        axis.text.x = element_text(size=7),
        legend.position="none")
f2
ggsave(f2, file = "../../figs/Westwide-PLUS_BurnedArea_byIncident_1999to2020.png",
       width=4, height=2.5, dpi = 300) # adjust dpi accordingly
```

Burned area (Option 2).

```{r fig.width=4, fig.height=2.5, warning=F}
## Plot
f3 <- ics.west.plus %>% group_by(START_YEAR) %>%
  summarize(burned_area = sum(FINAL_ACRES)) %>%
  ungroup() %>%
  ggplot(aes(x=START_YEAR, y=burned_area)) +
  geom_line() +
  geom_point(aes(size=burned_area, fill=burned_area), shape=21, color = "gray20") +
  scale_fill_viridis_c(option="rocket", trans="sqrt") +
  scale_size(1,5) +
  scale_y_continuous(labels = scales::label_number(suffix = " M", scale = 1e-6)) +
  scale_x_continuous(limits=c(1999, 2020)) +
  coord_cartesian(clip = "off") +
  labs(x="\nIgnition Year", y="Burned Area (acres)\n", title="") +
  theme_minimal() +
  theme(plot.title = element_text(size = 9),
        plot.subtitle = element_text(size=7),
        plot.caption = element_text(size=6),
        axis.title.x = element_text(size=8),
        axis.title.y = element_text(size=8),
        axis.text.y = element_text(size=7),
        axis.text.x = element_text(size=7),
        legend.position="none")
f3
ggsave(f3, file = "../../figs/Westwide-PLUS_BurnedArea_1999to2020.png",
       width=4, height=2.5, dpi = 300) # adjust dpi accordingly
```

```{r fig.height=5, fig.width=7.5, warning=F}
arr1 <- ggarrange(ggarrange(f1, f3, nrow=2, ncol=1, align="v"), 
                  map1, ncol=2, nrow=1, align="h")
# arr1 <- annotate_figure(
#   arr1, bottom = text_grob("Data Source: St. Denis et al. (2020), updated", 
#                            face="italic", color = "black",
#                            hjust = 1, x = 1, size = 8))
arr1
ggsave(arr1, file = "../../figs/Westwide_HomeLoss_Burned_Area_byIncident_wMap_1999to2020.png",
       width=7.5, height=5, dpi = 300) # adjust dpi accordingly
```


# Joining ICS-209-PLUS with FIRED (ICS+FIRED)

Read in the add-in records and tidy.

```{r}
addin <- read.csv("../../data/tabular/mod/wf-incidents/ics-fired_incident_addin.csv") %>%
  mutate_at(vars(contains("_DATE")), as.Date, '%m/%d/%Y') %>%
  rename(X = ï..X) %>%
  mutate(EVACUATION_REPORTED = as.character(EVACUATION_REPORTED),
         COMPLEX = as.character(COMPLEX)) %>%
  st_as_sf(., coords=c("LONGITUDE", "LATITUDE"), na.fail=TRUE) %>%
  st_set_crs(st_crs(4326)) %>%
  mutate(LONGITUDE = unlist(map(.$geometry,1)),
         LATITUDE = unlist(map(.$geometry,2))) %>%
  rename(geom = geometry) %>%
  st_transform(., crs=st_crs(ics.west.plus))

# Handle the Cold Spring Complex (WA, 2020)
ics.west.plus <- ics.west.plus %>%
  filter(!INCIDENT_ID == "2020_11903764_WHITNEY",
         !INCIDENT_ID == "2020_11902323_PEARL HILL",
         !INCIDENT_ID == "2020_11901760_COLD SPRINGS") %>%
  bind_rows(., addin)
```

By spatial buffer overlap.

```{r warning=F}
## Remove records reporting 0 acres (likely erroneous)
ics.west.plus <- ics.west.plus %>% 
  filter(FINAL_ACRES != 0,
         FINAL_KM2 != 0,
         START_YEAR >= 2001,
         !is.na(DISCOVERY_DATE))
## Buffer incident points by generous 50km
ics.buffer <- st_buffer(ics.west.plus, 50000)
## Join to any overlapping FIRED perimeters
buffer.join <- st_join(fired.west.plus, ics.buffer, join=st_intersects)
```

A 50km buffer join to all events yields over two million records. Start filtering them down by attributes.

Ignition date filters ...

```{r}
## By DOY
doy.filter <- buffer.join %>% 
  mutate(doy_diff = (abs(as.numeric(difftime(DISCOVERY_DATE, ig_date, units = "days"))))) %>%
  filter(START_YEAR == ig_year & doy_diff <= 20)
dim(doy.filter)[1]
```

Fire size filters ...

```{r}
area.filter <- doy.filter %>% 
  mutate(ar_abs_dif = abs(tot_ar_km2 - FINAL_KM2),
         ar_chg_dif = abs((tot_ar_km2 - FINAL_KM2)/FINAL_KM2) * 100) %>%
  group_by(INCIDENT_ID) %>%
  filter(ar_chg_dif == min(ar_chg_dif),
         ar_abs_dif == min(ar_abs_dif)) %>%
  ungroup() %>%
  group_by(id) %>%
  filter(ar_chg_dif == min(ar_chg_dif),
         ar_abs_dif == min(ar_abs_dif)) %>%
  ungroup()
dim(area.filter)[1]
```

Now double check on how many duplicates we still have.

```{r}
dim(area.filter %>% group_by(INCIDENT_ID) %>% filter(n()>1))[1]
dim(area.filter %>% group_by(id) %>% filter(n()>1))[1]
## Check for just destructive events
print("For destructive incidents (Homes Destroyed > 0): ...")
dim(area.filter%>%filter(STR_DESTROYED_RES_TOTAL>0)%>%group_by(INCIDENT_ID)%>%filter(n()>1))[1]
dim(area.filter%>%filter(STR_DESTROYED_RES_TOTAL>0)%>%group_by(id)%>%filter(n()>1))[1]
## And by fire size
print("For burned area (acres > 1000): ...")
dim(area.filter%>%filter(FINAL_ACRES>1000)%>%group_by(INCIDENT_ID)%>%filter(n()>1))[1]
dim(area.filter%>%filter(FINAL_ACRES>1000)%>%group_by(id)%>%filter(n()>1))[1]
```

For the most part, the duplicates represent smaller, less destructive fires. we can try to filter again by the minimum date difference.

```{r}
min.filter <- area.filter %>%
  group_by(INCIDENT_ID) %>%
  filter(doy_diff == min(doy_diff)) %>%
  ungroup() %>%
  group_by(id) %>%
  filter(doy_diff == min(doy_diff)) %>%
  ungroup()
dim(min.filter)[1]
```

Again, double check on how many duplicates we still have.

```{r}
dim(min.filter %>% group_by(INCIDENT_ID) %>% filter(n()>1))[1]
dim(min.filter%>% group_by(id) %>% filter(n()>1))[1]
## Check for just destructive events
print("For destructive incidents (Homes Destroyed > 0): ...")
dim(min.filter%>%filter(STR_DESTROYED_RES_TOTAL>0)%>%group_by(INCIDENT_ID)%>%filter(n()>1))[1]
dim(min.filter%>%filter(STR_DESTROYED_RES_TOTAL>0)%>%group_by(id)%>%filter(n()>1))[1]
## And by fire size
print("For burned area (acres > 1000): ...")
dim(min.filter%>%filter(FINAL_ACRES>1000)%>%group_by(INCIDENT_ID)%>%filter(n()>1))[1]
dim(min.filter%>%filter(FINAL_ACRES>1000)%>%group_by(id)%>%filter(n()>1))[1]
```

Try to eliminate any exact duplicate rows.

```{r}
min.filter <- min.filter %>% dplyr::select(-c(fid_, STUSPS.x, X, X.1)) %>%
  distinct(.keep_all = T)
dim(min.filter)[1]
sum(min.filter$STR_DESTROYED_RES_TOTAL) / sum(ics.west.plus$STR_DESTROYED_RES_TOTAL) * 100
```

Isolate duplicate records, export for inspection.

```{r}
d <- min.filter %>% 
  group_by(INCIDENT_ID) %>% 
  filter(n()>1) %>%
  ungroup()
dd <- min.filter %>% 
  group_by(id) %>% 
  filter(n()>1) %>%
  ungroup()
```

Isolate non duplicates.

```{r}
min.filter_ <- min.filter %>% 
  filter(!INCIDENT_ID %in% d$INCIDENT_ID,
         !id %in% dd$id)
dim(min.filter_ %>% group_by(INCIDENT_ID) %>% filter(n()>1))[1]
dim(min.filter_ %>% group_by(id) %>% filter(n()>1))[1]
sum(min.filter_$STR_DESTROYED_RES_TOTAL) / sum(ics.west.plus$STR_DESTROYED_RES_TOTAL) * 100
```

Calculate nearest neighbor for incidents with duplicates.

```{r}
## Calculate centroid for FIRED event
fired.sub <- fired.west.plus %>% filter(!id %in% min.filter_$id)
centroid <- st_centroid(fired.sub)
## Get the incidents that have duplicates
ics.sub <- ics.west.plus %>% filter(!INCIDENT_ID %in% min.filter_$INCIDENT_ID)
## Try a join
t <- st_join(ics.sub, centroid, join = nngeo::st_nn, k = 1, maxdist = 25000)
## Filter down
tt <- t %>%
  mutate(doy_diff = (abs(as.numeric(difftime(DISCOVERY_DATE, ig_date, units = "days"))))) %>%
  filter(START_YEAR == ig_year & doy_diff <= 20) %>% 
  mutate(ar_abs_dif = abs(tot_ar_km2 - FINAL_KM2),
         ar_chg_dif = abs((tot_ar_km2 - FINAL_KM2)/FINAL_KM2) * 100) %>%
  group_by(INCIDENT_ID) %>%
  filter(ar_chg_dif == min(ar_chg_dif),
         ar_abs_dif == min(ar_abs_dif)) %>%
  ungroup() %>%
  group_by(id) %>%
  filter(ar_chg_dif == min(ar_chg_dif),
         ar_abs_dif == min(ar_abs_dif)) %>%
  ungroup() %>%
  group_by(INCIDENT_ID) %>%
  filter(doy_diff == min(doy_diff)) %>%
  ungroup() %>%
  group_by(id) %>%
  filter(doy_diff == min(doy_diff)) %>%
  ungroup()
## At this point, remove all duplicates
tt <- tt %>% distinct(id, .keep_all = TRUE)
## Double check duplicates
dim(tt %>% group_by(INCIDENT_ID) %>% filter(n()>1))[1]
dim(tt %>% group_by(id) %>% filter(n()>1))[1]
```

```{r}
## Now grab FIRED perimeter,
tt.fired.df <- inner_join(fired.conus%>%as_tibble(), tt, by="id", left=F) %>% st_as_sf()
# Tidy it
tt.fired.df <- tt.fired.df %>% select(-contains(".x"))
## Join back to df
dim(min.filter_%>%filter(INCIDENT_ID %in% tt$INCIDENT_ID))
final.join <- bind_rows(min.filter_, tt.fired.df) %>%
  select(-c(X, X.1, doy_diff, ar_abs_dif, ar_chg_dif, bupr_sum, bupr_sum1k, geom.x),
         -contains(".y"))
  
## Final test of duplicates
dim(final.join %>% group_by(INCIDENT_ID) %>% filter(n()>1))[1]
dim(final.join %>% group_by(id) %>% filter(n()>1))[1]
```

Final stats.

```{r}
print(paste("Percent of homes destroyed by western wildfires:  ",
            sum(final.join$STR_DESTROYED_RES_TOTAL) / sum(ics.west.plus$STR_DESTROYED_RES_TOTAL) * 100, sep=""))
print(paste("Percent of homes destroyed by CONUS wildfires:  ",
            sum(final.join$STR_DESTROYED_RES_TOTAL) / sum(ics.conus$STR_DESTROYED_RES_TOTAL) * 100, sep=""))
print(paste("Percent of acres burned by western wildfires:  ",
            sum(final.join$FINAL_ACRES) / sum(ics.west.plus$FINAL_ACRES) * 100, sep=""))
print(paste("Percent of acres burned by CONUS wildfires:  ",
            sum(final.join$FINAL_ACRES) / sum(ics.conus$FINAL_ACRES) * 100, sep=""))
```

Attach BUPR summaries.

```{r}
# Read in the summary stats by FIRED id
bupr.sums <- read.csv("../../../fired/data/events/mod/fired_conus_bupr_sum.csv") %>%
  as_tibble() %>%
  dplyr::select(-c(X, bupr_sum.1))
# Join to df
final.join_ <- inner_join(final.join, bupr.sums, by="id")
glimpse(final.join_)
```

Re-create maps with joined data.

```{r fig.height=5, fig.width=5}
map2 <- ggplot() +
  geom_sf(data=final.join%>%filter(STR_DESTROYED_RES_TOTAL==0), color="gray60", size=0.3, alpha=0.9) +
  geom_sf(data=states%>%filter(STUSPS %in% states.sub), fill=NA, color="gray20") +
  geom_sf(data=final.join%>%filter(STR_DESTROYED_RES_TOTAL>0)%>%arrange(STR_DESTROYED_RES_TOTAL),
          aes(color=STR_DESTROYED_RES_TOTAL, size=FINAL_ACRES), alpha=0.8) +
  scale_size(range = c(1,6)) +
  scale_color_viridis_c(option="inferno", trans="log10") +
  labs(color = "Homes Destroyed") +
  coord_sf(crs = st_crs(ics.conus)) +
  theme_void() +
  guides(color = guide_colourbar(position="top", barwidth = 0.4, barheight = 6, ticks=F,
                                 label.position = "right", title.position = "left",
                                 label.theme = element_text(angle = 0, size = 8)),
         size="none") +
  theme(legend.title = element_text(angle = 90, size=10),
        legend.position=c(0.85, 0.7))
map2
ggsave(map2, file = paste("../../figs/Map1-ICS209_HomesDestroyed_Area_WestPlus_Joined.png", sep=""), 
       width=5, height=5, dpi = 350) # adjust dpi accordingly
```

```{r fig.width=4, fig.height=2.5, warning=F}
## reorder factors
final.join$INCIDENT_NAME <- 
  reorder(final.join$INCIDENT_NAME, final.join$STR_DESTROYED_RES_TOTAL)
final.join$INCIDENT_NAME <- 
  factor(final.join$INCIDENT_NAME, levels=rev(levels(final.join$INCIDENT_NAME)))
# Grab the top 15
top10 <- final.join %>% slice_max(STR_DESTROYED_RES_TOTAL, n=10)
# Plot and label
f4 <- ggplot(data=final.join, aes(x=START_YEAR, y=STR_DESTROYED_RES_TOTAL, 
                                     group=factor(INCIDENT_NAME), 
                                     fill=STR_DESTROYED_RES_TOTAL), color="grey90") +
  geom_bar(stat="identity", position='stack') +
  scale_fill_viridis_c(option="inferno", trans="log10") +
  labs(
    x="", y="Homes Destroyed\n", fill="") +
    # title="Homes destroyed by Western wildfires (1999-2020)\n",
    # caption="Maxwell C. Cook, PhD Student, Department of Geography,
    # University of Colorado, Boulder\nData source: St. Denis et al., 2020 (updated)") +
  # ggrepel::geom_text_repel(data=. %>% filter(INCIDENT_ID %in% top10$INCIDENT_ID), 
  #                          aes(label=INCIDENT_NAME), position=position_jitterdodge(0.5), size = 2) +
  theme_minimal() +
  theme(plot.title = element_text(size = 9),
        plot.subtitle = element_text(size=7),
        plot.caption = element_text(size=6),
        axis.title.x = element_text(size=8),
        axis.title.y = element_text(size=8),
        axis.text.y = element_text(size=7),
        axis.text.x = element_text(size=7),
        legend.position="none")
f4
ggsave(f4, file = "../../figs/Westwide-PLUS_HomeLoss_byIncident_1999to2020_Joined.png",
       width=4, height=2.5, dpi = 300) # adjust dpi accordingly
rm(top10)
```

```{r fig.width=4, fig.height=2.5, warning=F}
## Plot
f5 <- final.join %>% group_by(START_YEAR) %>%
  summarize(burned_area = sum(FINAL_ACRES)) %>%
  ungroup() %>%
  ggplot(aes(x=START_YEAR, y=burned_area)) +
  geom_line() +
  geom_point(aes(size=burned_area, fill=burned_area), shape=21, color = "gray20") +
  scale_fill_viridis_c(option="rocket", trans="sqrt") +
  scale_size(1,5) +
  scale_y_continuous(labels = scales::label_number(suffix = " M", scale = 1e-6)) +
  scale_x_continuous(limits=c(1999, 2020)) +
  coord_cartesian(clip = "off") +
  labs(x="\nIgnition Year", y="Burned Area (acres)\n", title="") +
  theme_minimal() +
  theme(plot.title = element_text(size = 9),
        plot.subtitle = element_text(size=7),
        plot.caption = element_text(size=6),
        axis.title.x = element_text(size=8),
        axis.title.y = element_text(size=8),
        axis.text.y = element_text(size=7),
        axis.text.x = element_text(size=7),
        legend.position="none")
f5
ggsave(f5, file = "../../figs/Westwide-PLUS_BurnedArea_1999to2020_Joined.png",
       width=4, height=2.5, dpi = 300) # adjust dpi accordingly
```

```{r fig.height=5, fig.width=7.5, warning=F}
arr2 <- ggarrange(ggarrange(f4, f5, nrow=2, ncol=1, align="v"), 
                  map2, ncol=2, nrow=1, align="h")
# arr1 <- annotate_figure(
#   arr1, bottom = text_grob("Data Source: St. Denis et al. (2020), updated", 
#                            face="italic", color = "black",
#                            hjust = 1, x = 1, size = 8))
arr2
ggsave(arr1, file = "../../figs/Westwide_HomeLoss_Burned_Area_byIncident_wMap_1999to2020_Joined.png",
       width=7.5, height=5, dpi = 300) # adjust dpi accordingly
```

Save the data.

```{r}
st_write(final.join_, "../../data/spatial/mod/ics-fired/ics-fired_west_plus_2001to2020.gpkg", 
         delete_dsn = TRUE)
```

Supplemental figure: home loss across all of CONUS.

```{r fig.height=5, fig.width=7.5}
map3 <- ggplot() +
  geom_sf(data=ics.conus%>%filter(STR_DESTROYED_RES_TOTAL==0), color="gray60", size=0.3, alpha=0.9) +
  geom_sf(data=states, fill=NA, color="gray20") +
  geom_sf(data=ics.conus%>%filter(STR_DESTROYED_RES_TOTAL>0)%>%arrange(STR_DESTROYED_RES_TOTAL),
          aes(color=STR_DESTROYED_RES_TOTAL, size=FINAL_ACRES), alpha=0.8) +
  scale_size(range = c(1,6)) +
  scale_color_viridis_c(option="inferno", trans="log10") +
  labs(color = "Homes Destroyed") +
  coord_sf(crs = st_crs(ics.conus)) +
  theme_void() +
  guides(color = guide_colourbar(position="top", barwidth = 0.4, barheight = 6, ticks=F,
                                 label.position = "right", title.position = "left",
                                 label.theme = element_text(angle = 0, size = 8)),
         size="none") +
  theme(legend.title = element_text(angle = 90, size=10),
        legend.position=c(0.9, 0.25))
map3
ggsave(map3, file = paste("../../figs/Map1-ICS209_HomesDestroyed_Area_CONUS_Joined.png", sep=""), 
       width=7.5, height=5, dpi = 350) # adjust dpi accordingly
```

Now join the remaining incidents (CONUS East).

```{r}
# Isolate eastern wildfires
ics.east <- ics.conus %>% filter(!STUSPS %in% west.plus)
fired.east <- fired.conus %>% filter(!STUSPS %in% west.plus)
```

By spatial buffer overlap.

```{r warning=F}
## Remove records reporting 0 acres (likely erroneous)
ics.east <- ics.east %>% 
  filter(FINAL_ACRES != 0,
         FINAL_KM2 != 0,
         START_YEAR >= 2001,
         !is.na(DISCOVERY_DATE))
## Buffer incident points by generous 50km
ics.buffer <- st_buffer(ics.east, 50000)
## Join to any overlapping FIRED perimeters
buffer.join <- st_join(fired.east, ics.buffer, join=st_intersects)
```

A 50km buffer join to all events yields over two million records. Start filtering them down by attributes.

Ignition date filters ...

```{r}
## By DOY
doy.filter <- buffer.join %>% 
  mutate(doy_diff = (abs(as.numeric(difftime(DISCOVERY_DATE, ig_date, units = "days"))))) %>%
  filter(START_YEAR == ig_year & doy_diff <= 20)
dim(doy.filter)[1]
```

Fire size filters ...

```{r}
area.filter <- doy.filter %>% 
  mutate(ar_abs_dif = abs(tot_ar_km2 - FINAL_KM2),
         ar_chg_dif = abs((tot_ar_km2 - FINAL_KM2)/FINAL_KM2) * 100) %>%
  group_by(INCIDENT_ID) %>%
  filter(ar_chg_dif == min(ar_chg_dif),
         ar_abs_dif == min(ar_abs_dif)) %>%
  ungroup() %>%
  group_by(id) %>%
  filter(ar_chg_dif == min(ar_chg_dif),
         ar_abs_dif == min(ar_abs_dif)) %>%
  ungroup()
dim(area.filter)[1]
```

Now double check on how many duplicates we still have.

```{r}
dim(area.filter %>% group_by(INCIDENT_ID) %>% filter(n()>1))[1]
dim(area.filter %>% group_by(id) %>% filter(n()>1))[1]
## Check for just destructive events
print("For destructive incidents (Homes Destroyed > 0): ...")
dim(area.filter%>%filter(STR_DESTROYED_RES_TOTAL>0)%>%group_by(INCIDENT_ID)%>%filter(n()>1))[1]
dim(area.filter%>%filter(STR_DESTROYED_RES_TOTAL>0)%>%group_by(id)%>%filter(n()>1))[1]
## And by fire size
print("For burned area (acres > 100): ...")
dim(area.filter%>%filter(FINAL_ACRES>100)%>%group_by(INCIDENT_ID)%>%filter(n()>1))[1]
dim(area.filter%>%filter(FINAL_ACRES>100)%>%group_by(id)%>%filter(n()>1))[1]
```

For the most part, the duplicates represent smaller, less destructive fires. we can try to filter again by the minimum date difference.

```{r}
min.filter <- area.filter %>%
  group_by(INCIDENT_ID) %>%
  filter(doy_diff == min(doy_diff)) %>%
  ungroup() %>%
  group_by(id) %>%
  filter(doy_diff == min(doy_diff)) %>%
  ungroup()
dim(min.filter)[1]
```

Again, double check on how many duplicates we still have.

```{r}
dim(min.filter %>% group_by(INCIDENT_ID) %>% filter(n()>1))[1]
dim(min.filter%>% group_by(id) %>% filter(n()>1))[1]
## Check for just destructive events
print("For destructive incidents (Homes Destroyed > 0): ...")
dim(min.filter%>%filter(STR_DESTROYED_RES_TOTAL>0)%>%group_by(INCIDENT_ID)%>%filter(n()>1))[1]
dim(min.filter%>%filter(STR_DESTROYED_RES_TOTAL>0)%>%group_by(id)%>%filter(n()>1))[1]
## And by fire size
print("For burned area (acres > 100): ...")
dim(min.filter%>%filter(FINAL_ACRES>100)%>%group_by(INCIDENT_ID)%>%filter(n()>1))[1]
dim(min.filter%>%filter(FINAL_ACRES>100)%>%group_by(id)%>%filter(n()>1))[1]
```

Try to eliminate any exact duplicate rows.

```{r}
min.filter <- min.filter %>% dplyr::select(-c(fid_, STUSPS.x, X, X.1)) %>%
  distinct(.keep_all = T)
dim(min.filter)[1]
sum(min.filter$STR_DESTROYED_RES_TOTAL) / sum(ics.east$STR_DESTROYED_RES_TOTAL) * 100
```

Isolate duplicate records, export for inspection.

```{r}
d <- min.filter %>% 
  group_by(INCIDENT_ID) %>% 
  filter(n()>1) %>%
  ungroup()
dd <- min.filter %>% 
  group_by(id) %>% 
  filter(n()>1) %>%
  ungroup()
```

Isolate non duplicates.

```{r}
min.filter_ <- min.filter %>% 
  filter(!INCIDENT_ID %in% d$INCIDENT_ID,
         !id %in% dd$id)
dim(min.filter_ %>% group_by(INCIDENT_ID) %>% filter(n()>1))[1]
dim(min.filter_ %>% group_by(id) %>% filter(n()>1))[1]
sum(min.filter_$STR_DESTROYED_RES_TOTAL) / sum(ics.east$STR_DESTROYED_RES_TOTAL) * 100
```

Calculate nearest neighbor for incidents with duplicates.

```{r}
## Calculate centroid for FIRED event
fired.sub <- fired.east %>% filter(!id %in% min.filter_$id)
centroid <- st_centroid(fired.sub)
## Get the incidents that have duplicates
ics.sub <- ics.east %>% filter(!INCIDENT_ID %in% min.filter_$INCIDENT_ID)
## Try a join
t <- st_join(ics.sub, centroid, join = nngeo::st_nn, k = 1, maxdist = 25000)
## Filter down
tt <- t %>%
  mutate(doy_diff = (abs(as.numeric(difftime(DISCOVERY_DATE, ig_date, units = "days"))))) %>%
  filter(START_YEAR == ig_year & doy_diff <= 20) %>% 
  mutate(ar_abs_dif = abs(tot_ar_km2 - FINAL_KM2),
         ar_chg_dif = abs((tot_ar_km2 - FINAL_KM2)/FINAL_KM2) * 100) %>%
  group_by(INCIDENT_ID) %>%
  filter(ar_chg_dif == min(ar_chg_dif),
         ar_abs_dif == min(ar_abs_dif)) %>%
  ungroup() %>%
  group_by(id) %>%
  filter(ar_chg_dif == min(ar_chg_dif),
         ar_abs_dif == min(ar_abs_dif)) %>%
  ungroup() %>%
  group_by(INCIDENT_ID) %>%
  filter(doy_diff == min(doy_diff)) %>%
  ungroup() %>%
  group_by(id) %>%
  filter(doy_diff == min(doy_diff)) %>%
  ungroup()
## At this point, remove all duplicates
tt <- tt %>% distinct(id, .keep_all = TRUE)
## Double check duplicates
dim(tt %>% group_by(INCIDENT_ID) %>% filter(n()>1))[1]
dim(tt %>% group_by(id) %>% filter(n()>1))[1]
```

```{r}
## Now grab FIRED perimeter,
tt.fired.df <- inner_join(fired.conus%>%as_tibble(), tt, by="id", left=F) %>% st_as_sf()
# Tidy it
tt.fired.df <- tt.fired.df %>% select(-contains(".x"))
## Join back to df
dim(min.filter_%>%filter(INCIDENT_ID %in% tt$INCIDENT_ID))
final.join <- bind_rows(min.filter_, tt.fired.df) %>%
  select(-c(X, X.1, doy_diff, ar_abs_dif, ar_chg_dif, bupr_sum, bupr_sum1k, geom.x),
         -contains(".y"))
  
## Final test of duplicates
dim(final.join %>% group_by(INCIDENT_ID) %>% filter(n()>1))[1]
dim(final.join %>% group_by(id) %>% filter(n()>1))[1]
```

Export.

```{r}
st_write(final.join, "../../data/spatial/mod/ics-fired/ics-fired_east_2001to2020.gpkg", 
         delete_dsn = TRUE)
```

