---
title: "case-study"
author: "Maxwell C. Cook"
date: "5/17/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("setup.R")
```

```{r warning=F}

# Bring in ICS counties from Wildfire Spatio-temporal Database (Kathryn McConnell)
ics_counties <- read_csv("../../data/tabular/mod/ics-spatial/ics_spatial_counties_9.1.22.csv")

# Bring in counties database
counties <- tigris::counties(state = NULL, cb = TRUE, year = 2020, class = "sf") %>%
 st_transform(st_crs(lambert.prj))

# Summarize county data
counties.sum <- ics_counties %>%
  filter(!is.na(STATEFP)) %>%
  group_by(GEOID) %>%
  summarise(FIRE_COUNT = n(),
            STR_DESTROYED_TOTAL = sum(STR_DESTROYED_TOTAL, na.rm = TRUE),
            FOD_FINAL_ACRES = sum(FOD_FINAL_ACRES, na.rm = TRUE),
            STR_THREATENED = sum(STR_THREATENED_MAX, na.rm=TRUE),
            TOTAL_PERSONNEL_SUM = sum(TOTAL_PERSONNEL_SUM, na.rm = TRUE),
            WF_MAX_FSR = max(WF_MAX_FSR, na.rm = TRUE),
            WF_MEAN_FSR = mean(WF_MAX_FSR, na.rm=TRUE),
            PEAK_EVACUATIONS = sum(PEAK_EVACUATIONS, na.rm = TRUE),
            PROJECTED_COSTS = sum(PROJECTED_FINAL_IM_COST, na.rm=TRUE))
  
# Create spatial sf version of summarized ICS data
counties.sf <- inner_join(counties %>% as_tibble(), counties.sum, by = "GEOID") %>%
   st_as_sf() %>%
   st_transform(st_crs(lambert.prj))

```

# Figure X. Fire distributions from ICS-SPATIAL (CONUS+AK, Histogram).

Map fire count for CONUS.

```{r fig.width=5.75, fig.height = 4.25}

# Grab CONUS counties
counties.sf.conus <- counties.sf %>% filter(!STUSPS %in% c("AK", "PR", "HI"))
counties <- st_read("../../../data/boundaries/political/TIGER/tl19_us_counties_conus.gpkg") %>%
 st_transform(st_crs(lambert.prj))

# Plot log fire counts
county.map.conus <- ggplot() +
   geom_sf(data = counties, fill = "grey", color = "gray89", lwd = 0.2) +
   geom_sf(data = counties.sf.conus, aes(fill = FIRE_COUNT),
           color = "gray89", lwd = 0.2) +
   labs(title="(a) log(Fire Counts) by County (Continental United States)") +
   theme_void() + 
   scico::scale_fill_scico(
      palette = "lajolla", begin = 0, na.value = "grey",
      name = "log(Fire Count)", trans = "log10") +
   guides(fill = guide_colourbar(direction = "horizontal", barwidth = 6, barheight = 0.5, 
                                 ticks=F, title.position = "top")) +
   theme(legend.title = element_text(angle = 0, size=8),
         legend.text = element_text(size=7),
         legend.position=c(0.2, 0.1),
         plot.title = element_text(size=10, hjust=0.2))

county.map.conus

ggsave(county.map.conus, file = "../../figures/ics209plus_fired_FireCounts_CONUS.png",
       width=6.5, height=4.25, dpi = 500, bg="white")

rm(counties.sf.conus)

```

Alaska fire count map.

```{r fig.width = 4, fig.height = 3}

counties.sf.ak <- counties.sf %>% filter(STUSPS == "AK") %>%
 st_transform(st_crs("ESRI:102006"))
counties.ak <- tigris::counties(state = NULL, cb = TRUE, year = 2020, class = "sf") %>%
 st_transform(st_crs("ESRI:102006")) %>%
 filter(STUSPS == "AK")

# Plot log fire counts
county.map.ak <- ggplot() +
   geom_sf(data = counties.ak, fill = "grey", color = "gray89", lwd = 0.2) +
   geom_sf(data = counties.sf.ak, aes(fill = FIRE_COUNT),
           color = "gray89", lwd = 0.2) +
   labs(title="(b) log(Fire Counts) by County (Alaska)") +
   theme_void() + 
   theme(legend.title=element_text(size=8), 
         legend.text=element_text(size=7),
         legend.position = "bottom",
         plot.title = element_text(size=10, hjust=0.25)) +
   scico::scale_fill_scico(
      palette = "lajolla", begin = 0, na.value = "grey",
      name = "log(Fire Count)", trans = "log10") +
   guides(fill="none")

county.map.ak

ggsave(county.map.ak, file = "../../figures/ics209plus_fired_FireCounts_AK.png",
       width=3.25, height=2.75, dpi = 300, bg="white") # adjust dpi accordingly

rm(counties.sf.ak, counties.ak)

```

Histogram.

```{r fig.height=2.25, fig.width=3.25}

df <- incidents %>%
   mutate(
      HIST = if_else(START_YEAR <= 2001, "Hist1", "NA"),
      HIST = if_else(START_YEAR >= 2002 & START_YEAR <= 2013, "Hist2", HIST),
      HIST = if_else(START_YEAR >= 2014, "Current", HIST)
   ) %>%
   group_by(START_YEAR, HIST) %>%
   summarize(
      counts = n()) %>%
   ungroup() %>%
  # Now manually update the field for overlap
  mutate(counts = if_else(HIST=="Hist1"&START_YEAR==2001, 1032, as.double(counts)),
         counts = if_else(HIST=="Hist2"&START_YEAR==2002, 1008, as.double(counts))) %>%
  add_row(HIST="Hist1", START_YEAR=2002, counts=147) %>%
  add_row(HIST="Hist2", START_YEAR=2001, counts=152)

histo <- ggplot(data=df, aes(x=START_YEAR, y=counts, fill=as.factor(HIST))) +
   geom_bar(stat="identity", position=position_stack(reverse=TRUE)) +
   scale_fill_manual(breaks=c("Hist1","Hist2","Current"),
                     values=c("light blue", "dark red", "orange")) +
   labs(y="",x="",title="(c) Number of wildfire incidents by year",
        fill="") +
   theme_minimal() +
   theme(
      plot.title = element_text(size=10, hjust=0.1),
      axis.title = element_text(size=8),
      axis.text.x = element_text(size=7,angle=35),
      axis.text.y = element_text(size=7),
      legend.position = c(0.12, 0.85),
      legend.text = element_text(size=6),
      legend.key.size = unit(0.8,"line")
   )

histo

ggsave(histo, file = "../../figures/ics209plus_fired_FireCounts_Histogram.png",
       width=3, height=2, dpi = 300, bg="white") 

rm(df)

```

Combine.

```{r fig.height = 5.25, fig.width = 6.5}

combined <- ggpubr::ggarrange(
   county.map.conus, ggpubr::ggarrange(county.map.ak, histo, ncol=2), nrow=2
)
combined

ggsave(combined, file = "../../figures/ics209plus_fired_FireCounts_Panel.png",
       width=6.5, height=5.25, dpi = 500, bg="white") # adjust dpi accordingly

rm(combined)

```

# Figure N. Summary of Key Variables

```{r}

# Grab CONUS counties
counties.sf.conus <- counties.sf %>% filter(!STUSPS %in% c("AK", "PR", "HI"))
counties.conus <- st_read("../../../data/boundaries/political/TIGER/tl19_us_counties_conus.gpkg") %>%
 st_transform(st_crs(lambert.prj))

# Total acres
acres <- ggplot() +
  geom_sf(data = counties.conus, fill = "grey", color = "gray89", lwd = 0.2) +
  geom_sf(data = counties.sf.conus, 
          aes(fill = log(FOD_FINAL_ACRES)),
          color = "gray89", lwd = 0.2) +
  labs(title="(a) Burned area (acres)") +
  theme_void() + 
  scico::scale_fill_scico(
    palette = "lajolla", begin = 0, na.value = "grey",
    name = "log(Burned Acres)") +
  theme(legend.title=element_text(size=6), 
        legend.text=element_text(size=5),
        legend.position=c(0.2, 0.1),
        plot.title = element_text(size=7)) +
  guides(fill = guide_colourbar(direction = "horizontal",
                                label.position = "bottom",
                                title.position="top",
                                barwidth = 4.25, barheight = 0.6))
ggsave(acres, file = "../../figures/ics209plus_fired_KeyVars_Acres.png",
       width=3.25, height=2.75, dpi = 350, bg="white") 

# Max spread rate
fsr <- ggplot() +
  geom_sf(data = counties.conus, fill = "grey", color = "gray89", lwd = 0.2) +
  geom_sf(data = counties.sf.conus, 
          aes(fill = log(WF_MAX_FSR)),
          color = "gray89", lwd = 0.2) +
  labs(title = "(b) Max spread rate (acres/day)") +
  theme_void() + 
  scico::scale_fill_scico(palette = "lajolla",
                          begin = 0,
                          na.value = "grey",
                          name = "log(Max. Spread Rate)") +
  theme(legend.title=element_text(size=6), 
        legend.text=element_text(size=5),
        legend.position=c(0.2, 0.1),
        plot.title=element_text(size=7)) +
  guides(fill = guide_colourbar(direction = "horizontal",
                                label.position = "bottom",
                                title.position="top",
                                barwidth = 4.25, barheight = 0.6)) 
ggsave(fsr, file = "../../figures/ics209plus_fired_KeyVars_FSR.png",
       width=3.25, height=2.75, dpi = 350, bg="white")

# Threatened structures
threat <- ggplot() +
  geom_sf(data = counties.conus, fill = "grey", color = "gray89", lwd = 0.2) +
  geom_sf(data = counties.sf.conus, 
          aes(fill = log(STR_THREATENED)),
          color = "gray89", lwd = 0.2) +
  labs(title="(e) Total structures threatened") +
  theme_void() + 
  scico::scale_fill_scico(palette = "lajolla",
                          begin = 0,
                          na.value = "grey",
                          name = "log(Structures Threatened)") +
  theme(legend.title=element_text(size=6), 
        legend.text=element_text(size=5),
        legend.position=c(0.2, 0.1),
        plot.title=element_text(size=7)) +
  guides(fill = guide_colourbar(direction = "horizontal",
                                label.position = "bottom",
                                title.position="top",
                                barwidth = 4.25, barheight = 0.6)) 
ggsave(threat, file = "../../figures/ics209plus_fired_KeyVars_Threat.png",
       width=3.25, height=2.75, dpi = 350, bg="white")


# Structures destroyed (log)
struct <- ggplot() +
  geom_sf(data = counties.conus, fill = "grey", color = "gray89", lwd = 0.2) +
  geom_sf(data = counties.sf.conus, 
          aes(fill = log(STR_DESTROYED_TOTAL)),
          color = "gray89", lwd = 0.2) +
  labs(title="(f) Total structures destroyed") +
  theme_void() +
  scico::scale_fill_scico(palette = "lajolla",
                          begin = 0,
                          na.value = "grey",
                          name = "log(Structures Destroyed)") +
  theme(legend.title=element_text(size=6), 
        legend.text=element_text(size=5),
        legend.position=c(0.2, 0.1),
        plot.title=element_text(size=7)) +
  guides(fill = guide_colourbar(direction = "horizontal",
                                label.position = "bottom",
                                title.position="top",
                                barwidth = 4.25, barheight = 0.6)) 
ggsave(struct, file = "../../figures/ics209plus_fired_KeyVars_Destr.png",
       width=3.25, height=2.75, dpi = 350, bg="white") 


# Personnel costs
personnel <- ggplot() +
  geom_sf(data = counties.conus, fill = "grey", color = "gray89", lwd = 0.2) +
  geom_sf(data = counties.sf.conus, 
          aes(fill = log(TOTAL_PERSONNEL_SUM)),
          color = "gray89", lwd = 0.2) +
  labs(title="(c) Total assigned personnel") +
  theme_void() + 
  scico::scale_fill_scico(palette = "lajolla",
                          begin = 0,
                          na.value = "grey",
                          name = "log(Assigned Personnel)") +
  theme(legend.title=element_text(size=6), 
        legend.text=element_text(size=6),
        legend.position=c(0.2, 0.1),
        plot.title=element_text(size=7)) +
  guides(fill = guide_colourbar(direction = "horizontal",
                                label.position = "bottom",
                                title.position="top",
                                barwidth = 4.25, barheight = 0.6)) 
ggsave(personnel, file = "../../figures/ics209plus_fired_KeyVars_Personnel.png",
       width=3.25, height=2.75, dpi = 350, bg="white")

# Projected IM costs
cost <- ggplot() +
  geom_sf(data = counties.conus, fill = "grey", color = "gray89", lwd = 0.2) +
  geom_sf(data = counties.sf.conus, 
          aes(fill = log(PROJECTED_COSTS)),
          color = "gray89", lwd = 0.2) +
  labs(title="(c) Projected IM costs ($)") +
  theme_void() + 
  scico::scale_fill_scico(palette = "lajolla",
                          begin = 0,
                          na.value = "grey",
                          name = "log(Projected Costs)") +
  theme(legend.title=element_text(size=6), 
        legend.text=element_text(size=5),
        legend.position=c(0.2, 0.1),
        plot.title=element_text(size=7)) +
  guides(fill = guide_colourbar(direction = "horizontal",
                                label.position = "bottom",
                                title.position="top",
                                barwidth = 4.25, barheight = 0.6)) 
ggsave(cost, file = "../../figures/ics209plus_fired_KeyVars_ProjCost.png",
       width=3.25, height=2.75, dpi = 350, bg="white")


# Peak evacuations
evacuation <- ggplot() +
  geom_sf(data = counties.conus, fill = "grey", color = "gray89", lwd = 0.2) +
  geom_sf(data = counties.sf.conus, 
          aes(fill = log(PEAK_EVACUATIONS)),
          color = "gray89", lwd = 0.2) +
  labs(title="(d) Total evacuations (2014-2020)") +
  theme_void() + 
  scico::scale_fill_scico(palette = "lajolla",
                          begin = 0,
                          na.value = "grey",
                          name = "log(Total Evacuations)") +
  theme(legend.title=element_text(size=6), 
        legend.text=element_text(size=5),
        legend.position=c(0.2, 0.1),
        plot.title=element_text(size=7)) +
  guides(fill = guide_colourbar(direction = "horizontal",
                                label.position = "bottom",
                                title.position="top",
                                barwidth = 4.25, barheight = 0.6)) 
ggsave(evacuation, file = "../../figures/ics209plus_fired_KeyVars_Evac.png",
       width=3.25, height=2.75, dpi = 350, bg="white")

```
 
Arrange.

```{r fig.height=9, fig.width=7}

# Try to arrange with ggarrange
combined <- ggpubr::ggarrange(
  acres, fsr, cost, evacuation, threat, struct,
  nrow=3, ncol=2, align="hv"
) +
  theme(plot.margin = margin(0.2,0.2,0.2,0.2))
# combined

ggsave(combined, file = "../../figures/ics209plus_fired_KeyVars_6Panel.png",
       width=7.25, height=9.50, dpi = 500, bg="white") 

```

# Figure 4. 2017 Chetco Bar Fire Case Study

Load in the MCD64A1 burn date raster for 2017 and mask it to the MTBS footprint for the Chetco Bar fire.

```{r warning=F}

library(raster,quietly=T)

# MTBS footprint
footprint <- mtbs %>%
   filter(Incid_Name == "CHETCO BAR") %>%
   st_transform(st_crs(wgs.prj))

# Define the 'burndate' raster from MCD64
burndate <- raster::raster("../../../earth-lab/fast-fires/data/burndate/annual/mcd64a1_y2017_burndate.tif")
# Re-define the projection and mask to MTBS footprint
crs(burndate) <- wgs.prj
burndate <- burndate %>%
 raster::crop(footprint) %>%
 raster::mask(footprint)
 
plot(burndate)

```

Examine the range of burn day values.

```{r}

# Plot the histogram of burn date
hist(burndate,
     main = "Distribution of day of burn (DOY) for Chetco Bar Fire (2017)",
     xlab = "Day of Year", ylab = "Frequency",
     col = "red")

```

Tidy up the day of burn values (removing sparse extremes on the end). Convert to days since ignition.

```{r}

# Cut of the sparse values > 255
burndate[burndate>258] = NA
# Now grab the range of burn days
range.doy <- range(burndate[],na.rm=TRUE)

# Grab some vars for start and finish
start <- as.numeric(range.doy[1])
breaks <- abs(range.doy[1]-range.doy[2])

# Update raster values to get "days since ignition"
burndate <- burndate - start
# Grab the new range
range.dsi <- range(burndate[],na.rm=TRUE)

# Extract as DF
df <- as.data.frame(burndate, xy = TRUE) %>% 
 rename(burndate = mcd64a1_y2017_burndate)
hist(df$burndate, 
     main="Distribution of days since ignition for Chetco Bar Fire (2017)", 
     xlab="Day Since Ignition", col="red")

```

Create the burn date plot.

```{r}

# Create the POO from ICS and from FIRED
ics.poo <- ics.points %>%
   filter(INCIDENT_NAME == "Chetco Bar" & START_YEAR == 2017)
fired.poo <- st_read("../../../FIRED/data/spatial/raw/events/conus-ak_to2022_ig_points_wgs.gpkg") %>%
 filter(id=="136250")

# Plot the map with POO coords
map <- ggplot() +
 geom_raster(data = df, aes(x = x, y = y, fill=burndate)) +
 scale_fill_fermenter(n.breaks=8, palette="Reds", na.value="white") +
 geom_sf(data=footprint, fill="transparent", size=0.35) +
 geom_sf(data=ics.poo, aes(color="Point of Origin"),
         shape=23, size=3.5, fill="black", color="white", stroke=1.2) +
 # geom_sf(data=fired.poo, shape=15, size=2.75, color="black") +
 labs(title="(a) Chetco Bar Fire MTBS Perimeter and FIRED Daily Progression",
      fill="Days Since Ignition") +
 theme_void() +
 theme(legend.position="bottom",
       plot.title = element_text(hjust = 0.5)) +
 guides(fill = guide_colourbar(title.position="top", title.hjust = 0.5,
                               barwidth = 12, barheight = 0.3,
                               ticks=TRUE, draw.llim=FALSE)) +
 ggsn::scalebar(footprint, dist = 5, st.dist = 0.03, st.size=2, height=-0.01,
                dist_unit="km", model = 'WGS84', transform=TRUE,
                border.size = 0.5, location = "bottomleft") +
 theme(plot.title = element_text(size = 7, hjust=-0.1, vjust=-0.3),
       legend.title = element_text(size = 7),
       legend.text = element_text(size=7),
       plot.margin = margin(0.6,0.6,0.6,0.6, "cm")) 

map

ggsave(map, file = "../../figures/ics209plus_fired_case-study_ChetcoBar.png",
       width=3, height=3.5, dpi = 500, bg="white")

```

Now the sitrep plots.

```{r warning=F}

# Load the daily SITREPS
sitreps <- read_csv("../../data/tabular/raw/wf-incidents/ics-209-plus-2.0/ics209-plus-wf_sitreps_1999to2020.csv") %>%
 filter(INCIDENT_ID == "2017_7265708_CHETCO BAR")

# Create a Day/Month field
sitreps <- sitreps %>%
   mutate(REPORT_TO_DATE = as.Date(REPORT_TO_DATE))

# Simple Fire Spread Rate
p1 <- ggplot() +
 geom_line(data=sitreps, aes(x=REPORT_TO_DATE, y=WF_FSR), size = 0.25, color="gray20") +
 geom_point(data=sitreps, aes(x=REPORT_TO_DATE, y=WF_FSR), 
            shape=19, size = 0.3) +
 scale_y_continuous(labels = scales::label_number(suffix = " K", scale = 1e-3)) +
 scale_x_date(date_labels = "%b %d", date_breaks = "20 days", date_minor_breaks = "5 days") + 
 labs(y="Acres/Day", x="",
      title="(b) Simple Fire Spread Rate (acres/day)") +
 theme_classic() +
 theme(plot.title = element_text(size=6, vjust=0.2, hjust=0.2),
       axis.title = element_text(size = 6),
       axis.text.y = element_text(size=5),
       axis.text.x = element_text(size=5,angle=35))

# Burned Area
p2 <- ggplot() +
 geom_line(data=sitreps, aes(x=REPORT_TO_DATE, y=ACRES), size = 0.25, color="gray20") +
 geom_point(data=sitreps, aes(x=REPORT_TO_DATE, y=ACRES), 
            shape=19, size = 0.3) +
 scale_y_continuous(labels = scales::label_number(suffix = " K", scale = 1e-3)) +
 scale_x_date(date_labels = "%b %d", date_breaks = "20 days", date_minor_breaks = "5 days") + 
 labs(y="Acres", x="",
      title="(c) Burned Area (acres)") +
 theme_classic() +
 theme(plot.title = element_text(size=6, vjust=0.2, hjust=0.2),
       axis.title = element_text(size = 6),
       axis.text.y = element_text(size=5),
       axis.text.x = element_text(size=5,angle=35))

# Threatened Structures
p3 <- ggplot() +
 geom_line(data=sitreps, aes(x=REPORT_TO_DATE, y=STR_THREATENED), size = 0.25, color="gray40") +
 geom_point(data=sitreps, aes(x=REPORT_TO_DATE, y=STR_THREATENED), 
            shape=19, size = 0.3) +
 scale_y_continuous(labels = scales::label_number(suffix = " K", scale = 1e-3)) +
 scale_x_date(date_labels = "%b %d", date_breaks = "20 days", date_minor_breaks = "5 days") +
 labs(y="# Structures", x="",
      title="(d) Threatened Structures") +
 theme_classic() +
 theme(plot.title = element_text(size=6, vjust=0.2, hjust=0.2),
       axis.title = element_text(size = 6),
       axis.text.y = element_text(size=5),
       axis.text.x = element_text(size=5,angle=35))

# Estimated Costs
p4 <- ggplot() +
 geom_line(data=sitreps, aes(x=REPORT_TO_DATE, y=EST_IM_COST_TO_DATE), size = 0.25, color="gray40") +
 geom_point(data=sitreps, aes(x=REPORT_TO_DATE, y=EST_IM_COST_TO_DATE), 
            shape=19, size = 0.3) +
 scale_y_continuous(labels = scales::label_number(suffix = " M", scale = 1e-6)) +
 scale_x_date(date_labels = "%b %d", date_breaks = "20 days", date_minor_breaks = "5 days") +
 labs(y="Cost ($)", x="",
      title="(e) Estimated IM Costs ($)") +
 theme_classic() +
 theme(plot.title = element_text(size=6, vjust=0.2, hjust=0.2),
       axis.title = element_text(size = 6),
       axis.text.y = element_text(size=5),
       axis.text.x = element_text(size=5,angle=35))

# Destroyed Structures
p5 <- ggplot() +
 geom_line(data=sitreps, aes(x=REPORT_TO_DATE, y=STR_DESTROYED), size = 0.25, color="gray40") +
 geom_point(data=sitreps, aes(x=REPORT_TO_DATE, y=STR_DESTROYED),
            shape=19, size = 0.3) +
 scale_x_date(date_labels = "%b %d", date_breaks = "20 days", date_minor_breaks = "5 days") +
 # scale_y_continuous(labels = scales::label_number(suffix = " K", scale = 1e-3)) +
 labs(y="# Structures", x="",
      title="(f) Destroyed Structures") +
 theme_classic() +
 theme(plot.title = element_text(size=6, vjust=0.2, hjust=0.2),
       axis.title = element_text(size = 6),
       axis.text.y = element_text(size=5),
       axis.text.x = element_text(size=5,angle=35))

# Total Assigned Personnel
p6 <- ggplot() +
 geom_line(data=sitreps, aes(x=REPORT_TO_DATE, y=TOTAL_PERSONNEL), size = 0.25, color="gray40") +
 geom_point(data=sitreps, aes(x=REPORT_TO_DATE, y=TOTAL_PERSONNEL),
            shape=19, size = 0.3) +
 scale_x_date(date_labels = "%b %d", date_breaks = "20 days", date_minor_breaks = "5 days") +
 # scale_y_continuous(labels = scales::label_number(suffix = " K", scale = 1e-3)) +
 labs(y="# Personnel", x="",
      title="(g) Total Assigned Personnel") +
 theme_classic() +
 theme(plot.title = element_text(size=6, vjust=0.2, hjust=0.2),
       axis.title = element_text(size = 6),
       axis.text.y = element_text(size=5),
       axis.text.x = element_text(size=5,angle=35))

# Evacuations ordered
p7 <- ggplot() +
 geom_line(data=sitreps, aes(x=REPORT_TO_DATE, y=RPT_EVACUATIONS), size = 0.25, color="gray40") +
 geom_point(data=sitreps, aes(x=REPORT_TO_DATE, y=RPT_EVACUATIONS),
            shape=19, size = 0.3) +
 scale_x_date(date_labels = "%b %d", date_breaks = "20 days", date_minor_breaks = "5 days") +
 # scale_y_continuous(labels = scales::label_number(suffix = " K", scale = 1e-3)) +
 labs(y="# Evacuations", x="Report Date",
      title="(h) Evacuations Ordered") +
 theme_classic() +
 theme(plot.title = element_text(size=6, vjust=0.2, hjust=0.2),
       axis.title = element_text(size = 6),
       axis.text.y = element_text(size=5),
       axis.text.x = element_text(size=5,angle=35))

# Aerial resources
p8 <- ggplot() +
 geom_line(data=sitreps, aes(x=REPORT_TO_DATE, y=TOTAL_AERIAL), size = 0.25, color="gray40") +
 geom_point(data=sitreps, aes(x=REPORT_TO_DATE, y=TOTAL_AERIAL),
            shape=19, size = 0.3) +
 scale_x_date(date_labels = "%b %d", date_breaks = "20 days", date_minor_breaks = "5 days") +
 # scale_y_continuous(labels = scales::label_number(suffix = " K", scale = 1e-3)) +
 labs(y="# Aerial", x="Report Date",
      title="(i) Total Aerial Resources") +
 theme_classic() +
 theme(plot.title = element_text(size=6, vjust=0.2, hjust=0.2),
       axis.title = element_text(size = 6),
       axis.text.y = element_text(size=5),
       axis.text.x = element_text(size=5,angle=35))

# Create the panel figure
panel <- ggpubr::ggarrange(p1,p2,p3,p4,p5,p6,p7,p8, ncol=2,nrow=4, vjust=-0.3, align = "hv") +
   theme(plot.margin = margin(0.4,0.4,0.4,0.4, "cm"))
panel

# Save the panel figure
ggsave(panel, file = "../../figures/ics209plus_fired_case-study_ChetcoBar_SitReps.png",
       width=6, height=7.5, dpi = 500, bg="white")

```

Arrange the panel plot with the map.

```{r fig.width=8.5}

combined <- ggpubr::ggarrange(
   map, ggpubr::ggarrange(p1,p2,p3,p4,p5,p6,p7,p8, ncol=2,nrow=4, align = "hv")) +
   theme(plot.margin = margin(0.6,0.6,0.6,0.6, "cm"))
combined

ggsave(combined, file = "../../figures/ics209plus_fired_case-study_ChetcoBar_SitReps_wMap.png",
       width=8, height=5.5, dpi = 650, bg="white") 

# Tidy up
rm(p1,p2,p3,p4,p5,p6,p7,p8,map,sitreps,df,burndate,panel,combined,ics.poo,fired.poo,footprint)

```

